---
title: "Paso 2.2"
description: |
  Análisis de clases latentes exploratoria y comparativa con predictores   
date: "`r withr::with_locale(new = c('LC_TIME' = 'C'), code =format(Sys.time(),'%B %d, %Y'))`"
author: "Andrés González Santa Cruz"
output:
  distill::distill_article:
    code_folding: true
    fig_height: 6
    fig_width: 8
    theme: flatly
    toc: yes
    toc_depth: 5
    toc_float: yes
    output_dir: "docs"
  toc_float:
    collapsed: false
    smooth_scroll: true
---

```{css zoom-lib-src, echo = FALSE}
script src = "https://ajax.googleapis.com/ajax/libs/jquery/3.4.1/jquery.min.js"
```

```{js zoom-jquery, echo = FALSE}
 $(document).ready(function() {
    $('body').prepend('<div class=\"zoomDiv\"><img src=\"\" class=\"zoomImg\"></div>');
    // onClick function for all plots (img's)
    $('img:not(.zoomImg)').click(function() {
      $('.zoomImg').attr('src', $(this).attr('src')).css({width: '100%'});
      $('.zoomDiv').css({opacity: '1', width: 'auto', border: '1px solid white', borderRadius: '5px', position: 'fixed', top: '50%', left: '50%', marginRight: '-50%', transform: 'translate(-50%, -50%)', boxShadow: '0px 0px 50px #888888', zIndex: '50', overflow: 'auto', maxHeight: '100%'});
    });
    // onClick function for zoomImg
    $('img.zoomImg').click(function() {
      $('.zoomDiv').css({opacity: '0', width: '0%'}); 
    });
  });
  
```

```{css hideOutput-lib-src, echo = FALSE}
<script src="hideOutput.js"></script> 
```

```{js hideOutput, echo = FALSE}
$(document).ready(function() {    
	$chunks = $('.fold');    
	$chunks.each(function () {      // add button to source code chunks     
	if ( $(this).hasClass('s') ) {       
		$('pre.r', this).prepend("<div class=\"showopt\">Show Source</div><br style=\"line-height:22px;\"/>");
       		$('pre.r', this).children('code').attr('class', 'folded');     
       		}      // add button to output chunks     
		if ( $(this).hasClass('o') ) {       
			$('pre:not(.r)', this).has('code').prepend("<div class=\"showopt\">Show Output</div><br style=\"line-height:22px;\"/>");       
			$('pre:not(.r)', this).children('code:not(r)').addClass('folded');        // add button to plots       
			$(this).find('img').wrap('<pre class=\"plot\"></pre>');       
			$('pre.plot', this).prepend("<div class=\"showopt\">Show Plot</div><br style=\"line-height:22px;\"/>");       
			$('pre.plot', this).children('img').addClass('folded');      
			}   
});    // hide all chunks when document is loaded   
	$('.folded').css('display', 'none')    // function to toggle the visibility   
	$('.showopt').click(function() {     
			var label = $(this).html();     
			if (label.indexOf("Show") >= 0) {       
				$(this).html(label.replace("Show", "Hide"));     
			} else {
			  $(this).html(label.replace("Hide", "Show"));     
			}     
	$(this).siblings('code, img').slideToggle('fast', 'swing');   
	}); 
}); 

```

```{=html}
<style type="text/css">
.showopt {   
  background-color: #004c93;   color: #FFFFFF;    width: 100px;   height: 20px;   text-align: center;   vertical-align: middle !important;   float: right;   font-family: sans-serif;   border-radius: 8px; 
  }

.showopt:hover {     
        background-color: #dfe4f2;
        color: #004c93; 
        }  
pre.plot {   
        background-color: white !important; 
        } 
.tablelines table, .tablelines td, .tablelines th {
        border: 1px solid black;
        }

.centrado {
    text-align: center;
}

.table.center {
    margin-left:auto; 
    margin-right:auto;
  }

/* https://vivekjaiskumar.medium.com/css-is-and-not-selector-17c942ec83f :is()*/

/* Applies to outputs that are not code other than R*/

pre {
  overflow-x: auto !important;
}
pre code {
  word-wrap: normal !important;
  white-space: pre !important;
}
/*
pre:not(.sourceCode) { 
  white-space: nowrap !important;
}
*/
.sourceCode { /* Important gives precedence  */
  font-size: 10px !important;
  line-height: 50% !important;
}

body{ /* Normal  */
      text-align: justify;
  }

.superbigimage{
    overflow-y:scroll;
    height:350px;
    white-space: nowrap;
    overflow-x: auto; 
    width:100%;
}
.superbigimage img{
    overflow-y: scroll;
    overflow-x: hidden;
}

.message { color:#446C6E; font-family: monospace;font-size: 10px; line-height: 110%; font-weight: bold;}
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 5px; text-align: justify;}
div.red { background-color:#e6bab1; border-radius: 5px; padding: 5px; text-align: justify;}

.pandoc-table { /* Should add !important; but it seems no necessary  */
  margin-left:auto; /* To center */
  margin-right:auto;
  border-collapse: collapse;
  table-layout: auto;
  font-size: 11px;
  overflow-y: auto;
  max-height:450px !important;
  white-space: nowrap;
  overflow-x: auto; 
  width:450px;
}

.pandoc-table th {/* header */
text-align: center !important;
font-size: 10px;
padding: 0px;
}

.pandoc-table td {
text-align: left !important;
font-size: 9px;
padding: 0px;
}

.pandoc-table caption {
    text-align: left !important;
    font-size: 11px !important;
}

.controlly{
    overflow-y:scroll;
    height:350px;
    overflow-x: scroll; 
}


.controlly2{
    overflow-y:scroll;
    height:550px;
    overflow-x: scroll; 
}

</style>
```
```{=html}
<!-- We gotta do each function to hide code and outputs per section, by every ID, we gotta create a different function -->
<script>
function myFunction1() {
    var x = document.getElementById("myDIV");
    if (x.style.display === "none") {
        x.style.display = "block";
    } else {
        x.style.display = "none";
    }
}
</script>

<script>
function myFunction2() {
    var x = document.getElementById("myDIV2");
    if (x.style.display === "none") {
        x.style.display = "block";
    } else {
        x.style.display = "none";
    }
}
</script>
```


Cargamos los datos

```{r, echo=FALSE}
rm(list = ls());gc()
print(Sys.time())
load("data2_lca2_2023_04_26.RData")
```

Cargamos los paquetes

```{r, echo=FALSE, results="hide"}
knitr::opts_chunk$set(echo = TRUE)

if(!require(poLCA)){install.packages("poLCA")}
if(!require(poLCAParallel)){devtools::install_github("QMUL/poLCAParallel@package")}
if(!require(compareGroups)){install.packages("compareGroups")}
if(!require(parallel)){install.packages("parallel")}
if(!require(Hmisc)){install.packages("Hmisc")}
if(!require(tidyverse)){install.packages("tidyverse")}
try(if(!require(sjPlot)){install.packages("sjPlot")})
if(!require(emmeans)){install.packages("emmeans")}
if(!require(nnet)){install.packages("nnet")}
if(!require(here)){install.packages("here")}
if(!require(doParallel)){install.packages("doParallel")}
if(!require(progress)){install.packages("progress")}
if(!require(caret)){install.packages("caret")}
if(!require(rpart)){install.packages("rpart")}
if(!require(rpart.plot)){install.packages("rpart.plot")}
if(!require(partykit)){install.packages("partykit")}
if(!require(randomForest)){install.packages("randomForest")}
if(!require(ggcorrplot)){install.packages("ggcorrplot")}
if(!require(polycor)){install.packages("polycor")}
if(!require(tableone)){install.packages("tableone")}
if(!require(broom)){install.packages("broom")}
if(!require(plotly)){install.packages("plotly")}
if(!require(rsvg)){install.packages("rsvg")}
if(!require(DiagrammeRsvg)){install.packages("DiagrammeRsvg")}

#if(!require(poLCA)){githubinstall::gh_install_packages("poLCA", ref = github_pull("14"))}

#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
```

Definimos ciertas constantes

```{r}
clus_iter= 500 #500
n_thread <- parallel::detectCores()
nrep <- clus_iter # number of different initial values (could be n_thread too)
n_class_max <- 10 # maximum number of classes to investigate
n_bootstrap <- 100 #30 # 50 number of bootstrap samples
print(n_thread)
```

# Análisis de clases latentes

```{r model,eval=T, echo=T, paged.print=TRUE, fig.cap="Gráfico esquemático", fig.align='center'}                                 
library(DiagrammeR) #⋉
gr_lca2<-
DiagrammeR::grViz("
digraph flowchart {
    fontname='Comic Sans MS'

  # Nodes
  subgraph samelevel {
    CAUSAL [label = 'Causal',fontsize=10,shape = box]
    MUJER_REC [label = 'Sexo\n(mujer)',fontsize=10,shape = box]
    EDAD_MUJER_REC [label = 'Edad\nmujer',fontsize=10,shape = box]
    HITO1_EDAD_GEST_SEM_REC [label = 'Edad\nGestacional\nHito 1',fontsize=10,shape = box]
    MACROZONA [label = 'Macrozona',fontsize=10,shape = box]
    PAIS_ORIGEN_REC [label = 'País de\norigen',fontsize=10,shape = box]
    PREV_TRAMO_REC [label = 'Previsión y\ntramo',fontsize=10,shape = box]
    PUEBLO_ORIGINARIO_REC [label = 'Pueblo\noriginario',fontsize=10,shape = box]
    
  {rank=same; rankdir= 'TB'; CAUSAL MUJER_REC EDAD_MUJER_REC HITO1_EDAD_GEST_SEM_REC MACROZONA PAIS_ORIGEN_REC PREV_TRAMO_REC PUEBLO_ORIGINARIO_REC}
  }
  LCA [label= 'Seis\nclases\nlatentes', shape= circle, style=filled, color=lightgrey, fontsize=10]
  
  inter [label = 'Interupción\nembarazo',fontsize=10,shape = box, height=.00002] # set the position of the inter node pos='15,100'

  # Nodes
  subgraph {
   LCA ->  {CAUSAL MUJER_REC EDAD_MUJER_REC HITO1_EDAD_GEST_SEM_REC MACROZONA PAIS_ORIGEN_REC PREV_TRAMO_REC PUEBLO_ORIGINARIO_REC} [rank=same; rankdir= 'TB'] 
}
  subgraph {
   LCA -> inter [minlen=14] #minlen es necesario para correr arrowhead = none; 
  {rank=same; LCA inter [rankdir='LR']}; #; 
}
}")#, width = 1200, height = 900
#https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3733703/
#Cohort matching on a variable associated with both outcome and censoring
#Cohort matching on a confounder. We let A denote an exposure, Y denote an outcome, and C denote a confounder and matching variable. The variable S indicates whether an individual in the source population is selected for the matched study (1: selected, 0: not selected). See Section 2-7 for details.
#https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7064555/
DPI = 1200
WidthCM = 21
HeightCM = 8

gr_lca2 %>%
  export_svg %>% charToRaw %>% rsvg_pdf("_flowchart_lca_adj.pdf")

gr_lca2 %>% export_svg()%>%charToRaw %>% rsvg(width = WidthCM *(DPI/2.54), height = HeightCM *(DPI/2.54)) %>% png::writePNG("_flowchart_lca0_adj.png")

htmlwidgets::saveWidget(gr_lca2, "_flowchart_lca_adj.html")
webshot::webshot("_flowchart_lca_adj.html", "_flowchart_lca_adj.png",vwidth = 1200, vheight = 900,
        zoom = 2)
```


## Modelo alternativo


### Análisis ACL alt

```{r, lca-get, echo=FALSE}
#Biemer, P. P., & Wiesen, C. (2002). Measurement error evaluation of self-reported drug use: a latent class analysis of the US National Household Survey on Drug Abuse. Journal of the Royal Statistical Society: Series A (Statistics in Society), 165(1), 97–119. doi:10.1111/1467-985x.00612  
#lca_entropia(x="ppio", seed= 2125, k= 8, f= f_preds, dat= mydata_preds, nbr_repet= 30, na_rm= T)
#3
#<div style="border: 1px solid #ddd; padding: 5px; overflow-y: scroll; height:400px; overflow-x: scroll; width:100%">
# f is the selected variables
# dat is the data
# nb_var is the number of selected variables
# k is the number of latent class generated
# nbr_repet is the number of repetition to  
# reach the convergence of EM algorithm
# x es el código para las variables de los modelos
#seed es el numero random para las semillas. ej: 4345.
#Modo de calcular el mejor modelo.
#z_ # 
#2023-01-20
#https://github.com/QMUL/poLCAParallel/blob/master/exec/3_blrt.R
#0h s
f_adj<-cbind(CAUSAL, EDAD_MUJER_REC, PUEBLO_ORIGINARIO_REC, PAIS_ORIGEN_REC, HITO1_EDAD_GEST_SEM_REC, MACROZONA, PREV_TRAMO_REC)~ outcome

seed<-2125
old <- Sys.time()

require(progress)

set.seed(seed)
model_array_adj <- list()  # Initialize an empty list to store the results
pb <- progress_bar$new(total = n_class_max, message_class = "Running poLCA")

#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
for (k in 1:n_class_max) {
  nrep_int <- nrep  # Initialize nrep to a reasonable value
  while (nrep_int >= 1) {  # Try running poLCA with decreasing nrep until nrep reaches 1
    tryCatch({
      set.seed(seed)
      mod <- poLCAParallel::poLCA(
        f_adj, 
        mydata_preds3 %>% dplyr::mutate(outcome=ifelse(outcome==1,1,0)), # %>% janitor::tabyl(outcome)
        nclass = k, 
        nrep = nrep_int, 
        maxiter = 1e4,
        n.thread = 12,
        verbose = FALSE
      )
      model_array_adj[[k]] <- mod  # Store the result if no error occurs
      break  # Exit the loop if poLCA succeeds
    }, error = function(e) {
      message(paste("Error in poLCA for k =", k, ", nrep =", nrep_int, ":", conditionMessage(e)))
      nrep_int <- nrep_int / 2  # Reduce nrep by half if poLCA fails
    })
  }
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_  
  pb$tick()  # Increment the progress bar
  cat(sprintf("\r%d%% completed", round(k/n_class_max*100)))  # Print progress percentage
  Sys.sleep(.05)
}

pb$terminate()  # Close the progress bar
cat(': Done')  # Print "Done" message  
                    
model_array_adj_ppio<-model_array_adj

#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#BOOTSTRAP#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_

new_med<-(Sys.time())
paste0("The model took ",round(new_med-old,2)," until every LCA was computed")
```

Luego calculamos la razón de verosimilitud mediante remuestreo bootstrap (BLRT) entre los distintos modelos con el que asume una clase menos.

```{r,blrt, echo=FALSE}
# store p values for each nclass, 1 to n_class_max
# store 0 for 1 number of class, ie this says you cannot have zero number of
# classes
p_value_array_adj <- c(0)
# for all number of classes investigated:
#   - store the log likelihood ratio
#   - store all bootstrap samples log likelihoods ratios
fitted_log_ratio_array_adj <- rep(NaN, n_class_max)
bootstrap_log_ratio_array_adj <- list()

#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
# do the bootstrap likelihood ratio test for each number of classes
for (nclass in 2:n_class_max) {

  # get the null and alt models
  # these are models with one number of class differences
  null_model_adj <- model_array_adj_ppio[[nclass - 1]]
  alt_model_adj <- model_array_adj_ppio[[nclass]]

  # for each bootstrap sample, store the log likelihood ratio here
  bootstrap_results_adj <- poLCAParallel::blrt(
    null_model_adj, alt_model_adj,
    n_bootstrap, n_thread, nrep
  )

  # log likelihood ratio to compare the two models
  fitted_log_ratio_array_adj[nclass] <- bootstrap_results_adj[["fitted_log_ratio"]]
  # store the log likelihoods ratios for all bootstrap samples
  bootstrap_log_ratio_array_adj[[nclass]] <-
    bootstrap_results_adj[["bootstrap_log_ratio"]]
  # store the p value for this nclass
  p_value_array_adj <- c(p_value_array_adj, bootstrap_results_adj[["p_value"]])
  
  #progress bar
  cat(paste0(round(nclass / n_class_max * 100), '% completed'))
  Sys.sleep(.05)
  if (nclass == n_class_max) cat(': Done')
  else cat('\014')
}
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
new<-(Sys.time())
time_diff <- (Sys.time() - old)/60
paste0("The model took ",round(new-old,2)," minutes")

model_array_ppio2 <- model_array_adj
fitted_log_ratio_array_adj_ppio <- fitted_log_ratio_array_adj
bootstrap_log_ratio_array_adj_ppio <- bootstrap_log_ratio_array_adj
bootstrap_results_adj_ppio <- bootstrap_results_adj
p_value_array_adj_ppio <- p_value_array_adj

# Get the BIC values for all models in model_array_ppio2
bic_values_adj <- sapply(model_array_ppio2, function(model) model$bic)

# Identify the index of the model with the lowest BIC
best_model_index_adj <- which.min(bic_values_adj)

# Select the best model
LCA_best_model_adj_ppio <- model_array_ppio2[[best_model_index_adj]]
#####################################################################################################################################################################
#Within poLCA, parameter estimates are obtained by a procedure that repeatedly improves estimates.
#This is stopped when no further improvements are obtained, or until a maximum number of iterations is reached. The starting values are the values at which such repetitions were started. Increasing the number 4 R. ACHTERHOF ET AL.of iterations (cycles within each estimation) and setting more different starting values for each repetition results in a greater likelihood that the global (rather than local) maximum of the log-likelihood function (and thus, the best possible solution) is reached. The maximum number of iterations was chosen as 10.000, and 500 different sets of starting values were used (thus going beyond the recommendations by Linzer & Lewis, 2011; Oberski, 2016). As such, the influence of chance was minimized while the reproducibility of the results was maximized# 
                     
```



### Resultados ACL alt

Hicimos un gráfico de los resultados

```{r, echo=FALSE}

# Initialize an empty data frame
tab_ppio2 <- data.frame()##

# Loop through each model
for (i in 2:n_class_max) {
  skip_to_next <- FALSE

  # Get the model and the previous model
  mod2 <- model_array_ppio2[[i]]
  mod2_min1 <- model_array_ppio2[[(i-1)]]

  # Check if the model has valid predictions
  if (is.null(mod2$predclass)) {
    skip_to_next <- TRUE
  }

  # If the model has valid predictions, calculate the measures and add them to the data frame
  if (!skip_to_next) {
    # Number of latent classes
    mod2$C <- max(t(matrix(apply(mod2$y, 2, max))))
    # Number of manifest variables
    mod2$J <- ncol(mod2$y)
    # Total number of items
    mod2$I <- mod2$J * mod2$C
    # Degrees of freedom
    mod2$df <- mod2$C^mod2$I - mod2$npar - 1
    # Chi-square test
    mod2$Chisq.pvalue <- (1 - pchisq(mod2$Chisq, mod2$df))
    # AIC
    mod2$aic <- round(mod2$aic, 2)
    # BIC
    mod2$bic <- round(mod2$bic, 2)
    # Adjusted BIC n*=(n+2)/24 (https://github.com/dlinzer/poLCA/issues/10)
    mod2$aBIC <- round((-2 * mod2$llik) + (log((mod2$N+2)/24) * mod2$npar), 2) 
    # Conditional AIC
    mod2$cAIC <-  round((-2 * mod2$llik) + (2 * mod2$Nobs * log(mod2$N/mod2$Nobs)), 2)
    # approximate weight of evidence criterion  #https://jbds.isdsa.org/public/journals/1/html/v2n2/qiu/
    mod2$awe <-  round((-2 * mod2$llik) + (2 * mod2$npar * log(mod2$Nobs)+1.5), 2)    
    # Gsq: deviance
    mod2$Gsq
    # Likelihood ratio test
    mod2$Gsq.pvalue <- (1 - pchisq(mod2$Gsq, mod2$df))
    # Relative entropy
    mod2$RelEnt <- round(relative.entropy(mod2), 2)
    # Entropy R-squared
    mod2$EntR2 <- round(entropy.R2(mod2), 2)
    # Deviance change
    mod2$DevChange <- round(mod2_min1$Gsq - mod2$Gsq, 2)
    # Degrees of freedom change
    mod2$dfChange <- mod2_min1$resid.df - mod2$resid.df
    # P-value for deviance change
    mod2$pvalDevChange <- round(pchisq(mod2$DevChange, mod2$dfChange, lower.tail = FALSE), 4)
    mod2$BLRT <- round(fitted_log_ratio_array_adj_ppio[[i]],2)
    mod2$BLRT.pvalue <- p_value_array_adj_ppio[[i]]
    # Add the model index to the data frame
    mod2$ModelIndex <- i

    tab_ppio2 <- rbind.data.frame(tab_ppio2, t(data.matrix(mod2[c("llik", "Chisq", "Chisq.pvalue", "resid.df", "aic", "bic", "aBIC", "cAIC", "awe", "Gsq", "Gsq.pvalue", "RelEnt", "EntR2", "DevChange", "dfChange", "pvalDevChange", "ModelIndex","BLRT", "BLRT.pvalue")])))

    #  } else {}
  }
}

# identify the list-like columns
list_cols2 <- sapply(tab_ppio2, is.list)
# unlist the list-like columns
unlisted_cols2 <- lapply(tab_ppio2[list_cols2], unlist)
# bind the unlisted columns as a data frame
tab_ppio2 <- cbind(tab_ppio2[!list_cols2], do.call(cbind, unlisted_cols2))
#Erase rownames
rownames(tab_ppio2) <- NULL

manualcolors2 <- c('indianred1', 'cornflowerblue', 'gray50', 'darkolivegreen4', 'slateblue2', 
                  'firebrick4', 'goldenrod4')
levels2 <- c("llik", "Chisq", "Chisq.pvalue", "resid.df", "aic", "bic", "aBIC", "cAIC", "awe",
            "Gsq", "Gsq.pvalue", "RelEnt", "EntR2", "DevChange", "dfChange",
            "pvalDevChange", "BLRT", "BLRT.pvalue")
labels2 <- c('Log-Verosimilitud', 'Chi2', 'valor p Chi2', 'Grados de libertad', 
            'Criterio de Información\nde Akaike(AIC)', 'Criterio de Información\nBayesiano (BIC)', 'BIC Ajustado (SABIC)', "AIC Corregido",'Peso de evidencia aproximado(awe)','G-squared/Deviance', 'Valor p G-squared', 'Entropía Relativa', 'Entropía R2', 'Cambio en Deviance\n(con modelo previo)', 'Grados de libertad del cambio', 'valor p cambio deviance', 'BLRT', 'valor p BLRT')
fig_lca_fit2<- tab_ppio2 %>%
  dplyr::mutate_if(is.character, as.numeric) %>%  # convert character columns to numeric
  tidyr::pivot_longer(cols = -ModelIndex, #"evryone but index"
                       names_to = "indices", values_to = "value", values_drop_na = F) %>%
  dplyr::mutate(indices = factor(indices, levels = levels2, labels = labels2)) %>%
  dplyr::filter(grepl("(AIC|BIC|awe)",indices, ignore.case=T))%>%
	dplyr::mutate(ModelIndex= factor(ModelIndex, levels=2:n_class_max)) %>% 
  ggplot(aes(x = ModelIndex, y = value, group = indices, color = indices, linetype = indices)) +
  geom_line(size = 1.5) +
  scale_color_manual(values = manualcolors) +
  #scale_linetype_manual(values = c("solid", "dashed", "dotted")) +
  labs(x = "Número de clases", y = "Valor", color = "Medida", linetype = "Medida") +
  #facet_wrap(.~indices, scales = "free_y", nrow = 4, ncol = 1) +
  theme_bw()

fig_lca_fit2
ggsave("_fig2_comparison_adj.png",fig_lca_fit2, dpi=600)

```

Luego en una tabla

::: controlly2
```{r, echo=FALSE}
tab_ppio2 %>%#
  dplyr::select(ModelIndex, everything()) %>% 
    dplyr::mutate_if(is.character, as.numeric) %>%  # convert character columns to numeric
    knitr::kable(format="markdown", caption="Fit measures of models")
```
:::

Presentamos el modelo con mejor ajuste

::: controlly
```{r, echo=FALSE}
print(LCA_best_model_adj_ppio) #
```
:::


```{r, echo=FALSE}
print(Sys.time())
save.image("data2_lca2_adj.RData")
```

```{r, echo=FALSE}
require(tidyverse)
sesion_info <- devtools::session_info()
dplyr::select(
  tibble::as_tibble(sesion_info$packages),
  c(package, loadedversion, source)
) %>% 
  DT::datatable(filter = 'top', colnames = c('Row number' =1,'Variable' = 2, 'Percentage'= 3),
              caption = htmltools::tags$caption(
        style = 'caption-side: top; text-align: left;',
        '', htmltools::em('Packages')),
      options=list(
initComplete = htmlwidgets::JS(
        "function(settings, json) {",
        "$(this.api().tables().body()).css({
            'font-family': 'Helvetica Neue',
            'font-size': '50%', 
            'code-inline-font-size': '15%', 
            'white-space': 'nowrap',
            'line-height': '0.75em',
            'min-height': '0.5em'
            });",#;
        "}")))
```
