---
title: "Paso 2.11"
description: |
  Análisis de clases latentes exploratoria y comparativa sin predictores (glca), sin pueblo originario, pero con año y recategorización de edad mujer y semana gestacional hito 1   
date: "`r withr::with_locale(new = c('LC_TIME' = 'C'), code =format(Sys.time(),'%B %d, %Y'))`"
author: "Andrés González Santa Cruz"
output:
  distill::distill_article:
    code_folding: true
    fig_height: 6
    fig_width: 8
    theme: flatly
    toc: yes
    toc_depth: 5
    toc_float: yes
    output_dir: "docs"
  toc_float:
    collapsed: false
    smooth_scroll: true
---

```{css zoom-lib-src, echo = FALSE}
script src = "https://ajax.googleapis.com/ajax/libs/jquery/3.4.1/jquery.min.js"
```

```{js zoom-jquery, echo = FALSE}
 $(document).ready(function() {
    $('body').prepend('<div class=\"zoomDiv\"><img src=\"\" class=\"zoomImg\"></div>');
    // onClick function for all plots (img's)
    $('img:not(.zoomImg)').click(function() {
      $('.zoomImg').attr('src', $(this).attr('src')).css({width: '100%'});
      $('.zoomDiv').css({opacity: '1', width: 'auto', border: '1px solid white', borderRadius: '5px', position: 'fixed', top: '50%', left: '50%', marginRight: '-50%', transform: 'translate(-50%, -50%)', boxShadow: '0px 0px 50px #888888', zIndex: '50', overflow: 'auto', maxHeight: '100%'});
    });
    // onClick function for zoomImg
    $('img.zoomImg').click(function() {
      $('.zoomDiv').css({opacity: '0', width: '0%'}); 
    });
  });
  
```

```{css hideOutput-lib-src, echo = FALSE}
<script src="hideOutput.js"></script> 
```

```{js hideOutput, echo = FALSE}
$(document).ready(function() {    
	$chunks = $('.fold');    
	$chunks.each(function () {      // add button to source code chunks     
	if ( $(this).hasClass('s') ) {       
		$('pre.r', this).prepend("<div class=\"showopt\">Show Source</div><br style=\"line-height:22px;\"/>");
       		$('pre.r', this).children('code').attr('class', 'folded');     
       		}      // add button to output chunks     
		if ( $(this).hasClass('o') ) {       
			$('pre:not(.r)', this).has('code').prepend("<div class=\"showopt\">Show Output</div><br style=\"line-height:22px;\"/>");       
			$('pre:not(.r)', this).children('code:not(r)').addClass('folded');        // add button to plots       
			$(this).find('img').wrap('<pre class=\"plot\"></pre>');       
			$('pre.plot', this).prepend("<div class=\"showopt\">Show Plot</div><br style=\"line-height:22px;\"/>");       
			$('pre.plot', this).children('img').addClass('folded');      
			}   
});    // hide all chunks when document is loaded   
	$('.folded').css('display', 'none')    // function to toggle the visibility   
	$('.showopt').click(function() {     
			var label = $(this).html();     
			if (label.indexOf("Show") >= 0) {       
				$(this).html(label.replace("Show", "Hide"));     
			} else {
			  $(this).html(label.replace("Hide", "Show"));     
			}     
	$(this).siblings('code, img').slideToggle('fast', 'swing');   
	}); 
}); 

```

```{=html}
<style type="text/css">
.showopt {   
  background-color: #004c93;   color: #FFFFFF;    width: 100px;   height: 20px;   text-align: center;   vertical-align: middle !important;   float: right;   font-family: sans-serif;   border-radius: 8px; 
  }

.showopt:hover {     
        background-color: #dfe4f2;
        color: #004c93; 
        }  
pre.plot {   
        background-color: white !important; 
        } 
.tablelines table, .tablelines td, .tablelines th {
        border: 1px solid black;
        }

.centrado {
    text-align: center;
}

.table.center {
    margin-left:auto; 
    margin-right:auto;
  }

/* https://vivekjaiskumar.medium.com/css-is-and-not-selector-17c942ec83f :is()*/

/* Applies to outputs that are not code other than R*/

pre {
  overflow-x: auto !important;
}
pre code {
  word-wrap: normal !important;
  white-space: pre !important;
}
/*
pre:not(.sourceCode) { 
  white-space: nowrap !important;
}
*/
.sourceCode { /* Important gives precedence  */
  font-size: 10px !important;
  line-height: 50% !important;
}

body{ /* Normal  */
      text-align: justify;
  }

.superbigimage{
    overflow-y:scroll;
    height:350px;
    white-space: nowrap;
    overflow-x: auto; 
    width:100%;
}
.superbigimage img{
    overflow-y: scroll;
    overflow-x: hidden;
}

.message { color:#446C6E; font-family: monospace;font-size: 10px; line-height: 110%; font-weight: bold;}
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 5px; text-align: justify;}
div.red { background-color:#e6bab1; border-radius: 5px; padding: 5px; text-align: justify;}

.pandoc-table { /* Should add !important; but it seems no necessary  */
  margin-left:auto; /* To center */
  margin-right:auto;
  border-collapse: collapse;
  table-layout: auto;
  font-size: 11px;
  overflow-y: auto;
  max-height:450px !important;
  white-space: nowrap;
  overflow-x: auto; 
  width:450px;
}

.pandoc-table th {/* header */
text-align: center !important;
font-size: 10px;
padding: 0px;
}

.pandoc-table td {
text-align: left !important;
font-size: 9px;
padding: 0px;
}

.pandoc-table caption {
    text-align: left !important;
    font-size: 11px !important;
}

.controlly{
    overflow-y:scroll;
    height:350px;
    overflow-x: scroll; 
}


.controlly2{
    overflow-y:scroll;
    height:550px;
    overflow-x: scroll; 
}

</style>
```
```{=html}
<!-- We gotta do each function to hide code and outputs per section, by every ID, we gotta create a different function -->
<script>
function myFunction1() {
    var x = document.getElementById("myDIV");
    if (x.style.display === "none") {
        x.style.display = "block";
    } else {
        x.style.display = "none";
    }
}
</script>

<script>
function myFunction2() {
    var x = document.getElementById("myDIV2");
    if (x.style.display === "none") {
        x.style.display = "block";
    } else {
        x.style.display = "none";
    }
}
</script>
```


Cargamos los datos

```{r, echo=FALSE}
rm(list = ls());gc()
load("data2.RData")
```

Cargamos los paquetes

```{r, echo=FALSE, results="hide"}
knitr::opts_chunk$set(echo = TRUE)

if(!require(poLCA)){install.packages("poLCA")}
if(!require(poLCAParallel)){devtools::install_github("QMUL/poLCAParallel@package")}
if(!require(compareGroups)){install.packages("compareGroups")}
if(!require(parallel)){install.packages("parallel")}
if(!require(Hmisc)){install.packages("Hmisc")}
if(!require(tidyverse)){install.packages("tidyverse")}
try(if(!require(sjPlot)){install.packages("sjPlot")})
if(!require(emmeans)){install.packages("emmeans")}
if(!require(nnet)){install.packages("nnet")}
if(!require(here)){install.packages("here")}
if(!require(doParallel)){install.packages("doParallel")}
if(!require(progress)){install.packages("progress")}
if(!require(caret)){install.packages("caret")}
if(!require(rpart)){install.packages("rpart")}
if(!require(rpart.plot)){install.packages("rpart.plot")}
if(!require(partykit)){install.packages("partykit")}
if(!require(randomForest)){install.packages("randomForest")}
if(!require(ggcorrplot)){install.packages("ggcorrplot")}
if(!require(polycor)){install.packages("polycor")}
if(!require(tableone)){install.packages("tableone")}
if(!require(DiagrammeR)){install.packages("DiagrammeR")}
if(!require(rsvg)){install.packages("rsvg")}
if(!require(DiagrammeRsvg)){install.packages("DiagrammeRsvg")}
if(!require(webshot)){install.packages("webshot")}
if(!require(htmlwidgets)){install.packages("htmlwidgets")}
if(!require(glca)){install.packages("glca")}

#if(!require(poLCA)){githubinstall::gh_install_packages("poLCA", ref = github_pull("14"))}

#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
lca_dir<-here::here()
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#
#	
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:

tryNA <- function(x){
    x <- try(x)
    if(inherits(x,'try-error')) return(NA)
    x
}

#https://rdrr.io/github/hyunsooseol/snowRMM/src/R/lca.b.R
#https://github.com/dlinzer/poLCA/issues/7

#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#' Bivariate residuals for latent class models
#' 
#' Calculate the "bivariate residuals" (BVRs) between pairs of variables 
#' in a latent class model.
#' 
#' This function compares the model-implied (expected) counts in the crosstables
#' of all pairs of observed dependent variables to the observed counts. For each
#' pair, it calculates a "chi-square" statistic,
#' 
#' \deqn{\text{BVR} = \sum_{j, j'} \frac{(n_{jj'} - e_{jj'})^2}{e_{jj'}}},
#' 
#' where \eqn{n_{jj'}} are the observed counts for categories \eqn{j} and \eqn{j'} 
#' of the variables being crosstabulated, and \eqn{e_{jj'}} are
#' the expected counts under the latent class model. 
#' 
#' Note that the BVR does not follow an asymptotic chi-square distribution and
#' for accurate p-values, parametric bootstrapping is necessary (Oberski et al. 2013).
#' 
#' @param fit A poLCA fit object
#' @param tol Optional: tolerance for small expected counts
#' @param rescale_to_df Optional: whether to divide the pairwise "chi-square" values by 
#' the degrees of freedom of the local crosstable. Default is TRUE.
#' @return The table of bivariate residuals
#' @author Daniel Oberski (daniel.oberski@gmail.com)
#' @seealso \code{\link{poLCA}} for fitting the latent class model.
#' @references 
#' Oberski, DL, Van Kollenburg, GH and Vermunt, JK (2013). 
#'   A Monte Carlo evaluation of three methods to detect local dependence in binary data latent class models. 
#'   Advances in Data Analysis and Classification 7 (3), 267-279.
#' @examples
#' data(values)
#' f <- cbind(A, B, C, D) ~ 1
#' M0 <- poLCA(f,values, nclass=1, verbose = FALSE) 
#' bvr(M0) # 12.4, 5.7, 8.3, 15.6, ... 
bvr <- function(fit, tol = 1e-3, rescale_to_df = TRUE) {
  stopifnot(class(fit) == "poLCA")

  ov_names <- names(fit$predcell)[1:(ncol(fit$predcell) - 2)]
  ov_combn <- combn(ov_names, 2)

  get_bvr <- function(ov_pair) {
    form_obs <- as.formula(paste0("observed ~ ", ov_pair[1], " + ", ov_pair[2]))
    form_exp <- as.formula(paste0("expected ~ ", ov_pair[1], " + ", ov_pair[2]))

    counts_obs <- xtabs(form_obs, data = fit$predcell)
    counts_exp <- xtabs(form_exp, data = fit$predcell)
    counts_exp <- ifelse(counts_exp < tol, tol, counts_exp) # Prevent Inf/NaN

    bvr_df <- prod(dim(counts_exp) - 1)
    bvr_value <- sum((counts_obs - counts_exp)^2 / counts_exp)

    if(rescale_to_df) bvr_value <- bvr_value / bvr_df

    attr(bvr_value, "df") <- bvr_df

    bvr_value
  }

  bvr_pairs <- apply(ov_combn, 2, get_bvr)

  attr(bvr_pairs, "rescale_to_df") <- rescale_to_df
  attr(bvr_pairs, "class") <- "dist"
  attr(bvr_pairs, "Size") <- length(ov_names)
  attr(bvr_pairs, "Labels") <- ov_names
  attr(bvr_pairs, "Diag") <- FALSE
  attr(bvr_pairs, "Upper") <- FALSE

  bvr_pairs
}

#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
poLCA.entropy.fix <- function (lc)
{
  K.j <- sapply(lc$probs, ncol)
  fullcell <- expand.grid(sapply(K.j, seq, from = 1))
  P.c <- poLCA.predcell(lc, fullcell)
  return(-sum(P.c * log(P.c), na.rm = TRUE))
}

#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#Calculate entropy R2 for poLCA model

# MIT license
# Author: Daniel Oberski
# Input: result of a poLCA model fit
# Output: entropy R^2 statistic (Vermunt & Magidson, 2013, p. 71)
# See: daob.nl/wp-content/uploads/2015/07/ESRA-course-slides.pdf
# And: https://www.statisticalinnovations.com/wp-content/uploads/LGtecnical.pdf
machine_tolerance <- sqrt(.Machine$double.eps)
entropy.R2 <- function(fit) {
  entropy <- function(p) {
    p <- p[p > machine_tolerance] # since Lim_{p->0} p log(p) = 0
    sum(-p * log(p))
  }
  error_prior <- entropy(fit$P) # Class proportions
  error_post <- mean(apply(fit$posterior, 1, entropy))
  R2_entropy <- (error_prior - error_post) / error_prior
  R2_entropy
}

#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#http://researchdata.gla.ac.uk/879/1/Survey_data_processed_using_R.pdf
##Function to plot variable probabilites by latent class

## Function to undertake chisquare analayis and plot graphs of residuals and contributions
chisquaretest.predictions.function <-
 function(indfactor.data,
	 predclass.data,
	 noclasses,
	 pitem,
	 gitem,
	 chirows,
	 chicols) {
	 chisquare.results <- chisq.test(indfactor.data, predclass.data)
	 residuals.data <- chisquare.results$residuals
	 colnames(residuals.data) <- chicols
	 rownames(residuals.data) <- chirows
		 title.text <-
			 paste(
			 "Residuals: chi Square Crosstabulation of\n",
			 pitem,
			 "and",
			 gitem,
			 "\n(Chisquare =",
			 round(chisquare.results$statistic, 3),
			 " p <",
			 round(chisquare.results$p.value, 3),
			 ")",
			 sep = " "
			 )
		 corrplot(
			 residuals.data,
			 is.cor = FALSE,
			 title = title.text,
			 mar = c(0, 0, 4, 0)
			 )
		 contrib.data <-
		 100 * residuals.data ^ 2 / chisquare.results$statistic
		 round(contrib.data, 3)
		 colnames(contrib.data) <- chicols
		 rownames(contrib.data) <- chirows
		 title.text <-
		 paste(
			 "Contributions: chi Square Crosstabulation of\n",
			 pitem,
			 "and",
			 gitem,
			 "\n(Chisquare =",
			 round(chisquare.results$statistic, 3),
			 " p <",
			 round(chisquare.results$p.value, 3),
			 ")",
			 sep = " "
			 )
		 corrplot(
			 contrib.data,
			 is.cor = FALSE,
			 title = title.text,
			 mar = c(0, 0, 4, 0)
			 )
		 return(chisquare.results)
 }
##Funciton for Cramers V test
cv.test = function(x, y) {
	 CV = sqrt(chisq.test(x, y, correct = FALSE)$statistic /
	 (length(x) * (min(
	 length(unique(x)), length(unique(y))
	 ) - 1)))
	 print.noquote("Cramér V / Phi:")
	 return(as.numeric(CV))
	}

#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:

if(.Platform$OS.type == "windows") withAutoprint({
  memory.size()
  memory.size(TRUE)
  memory.limit(size=56000)
})

path<-try(dirname(rstudioapi::getSourceEditorContext()$path))

options(knitr.kable.NA = '')


#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:

knitr::knit_hooks$set(time_it = local({
  now <- NULL
  function(before, options) {
    if (before) {
      # record the current time before each chunk
      now <<- Sys.time()
    } else {
      # calculate the time difference after a chunk
      res <- ifelse(difftime(Sys.time(), now)>(60^2),difftime(Sys.time(), now)/(60^2),difftime(Sys.time(), now)/(60^1))
      # return a character string to show the time
      x<-ifelse(difftime(Sys.time(), now)>(60^2),paste("Time for this code chunk to run:", round(res,1), "hours"),paste("Time for this code chunk to run:", round(res,1), "minutes"))
      paste('<div class="message">', gsub('##', '\n', x),'</div>', sep = '\n')
    }
  }
}))
knitr::opts_chunk$set(time_it = TRUE)
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:

#to format rows in bold
format_cells <- function(df, rows ,cols, value = c("italics", "bold", "strikethrough")){

  # select the correct markup
  # one * for italics, two ** for bold
  map <- setNames(c("*", "**", "~~"), c("italics", "bold", "strikethrough"))
  markup <- map[value]  

  for (r in rows){
    for(c in cols){

      # Make sure values are not factors
      df[[c]] <- as.character( df[[c]])

      # Update formatting
      df[r, c] <- ifelse(nchar(df[r, c])==0,"",paste0(markup, gsub(" ", "", df[r, c]), markup))
    }
  }

  return(df)
}
#To produce line breaks in messages and warnings
knitr::knit_hooks$set(
   error = function(x, options) {
     paste('\n\n<div class="alert alert-danger">',
           gsub('##', '\n', gsub('^##\ Error', '**Error**', x)),
           '</div>', sep = '\n')
   },
   warning = function(x, options) {
     paste('\n\n<div class="alert alert-warning">',
           gsub('##', '\n', gsub('^##\ Warning:', '**Warning**', x)),
           '</div>', sep = '\n')
   },
   message = function(x, options) {
     paste('<div class="message">',
           gsub('##', '\n', x),
           '</div>', sep = '\n')
   }
)
#
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:

as.data.frame.TableOne <- function(x, ...) {capture.output(print(x,
                          showAllLevels = TRUE, ...) -> x)
  y <- as.data.frame(x)
  y$charactersitic <- dplyr::na_if(rownames(x), "")
  y <- y %>%
  fill(charactersitic, .direction = "down") %>%
  select(charactersitic, everything())
  rownames(y) <- NULL
  y}
```

Definimos ciertas constantes

```{r}
clus_iter= 500#500 #500
n_thread <- parallel::detectCores()
nrep <- clus_iter # number of different initial values (could be n_thread too)
n_class_max <- 12 # maximum number of classes to investigate
n_bootstrap <- 100#00 #30 # 50 number of bootstrap samples
print(n_thread)
```

# Análisis de clases latentes

## Especificar modelo

Los valores ceros se declararon datos perdidos.

```{r, setup, echo=FALSE}
#table(data2$MACROZONA,data2$REGION_RESIDENCIA, exclude=NULL)
preds <- c("AÑO","CAUSAL","EDAD_MUJER_REC", "PUEBLO_ORIGINARIO_REC","PAIS_ORIGEN_REC", "HITO1_EDAD_GEST_SEM_REC", "MACROZONA", "PREV_TRAMO_REC")

#c("AÑO","MES_NUM",    "HITO3_COMPLICACION_POST_IVE", "HITO3_CONDICION_MUJER_POST_IVE", "HITO3_TIPO_ATENCION",  "HITO3_SERV_SALUD_ESTABLECIMIENTO", "HITO4_MUJER_ACEPTA_ACOM"),

mydata_preds <- data2%>% 
  #18-24 es el 18%, entre 25 y 35 es el 42% y el resto q es 30% som los mayores a 35
  #  NA=1;  >=4 & <14 =2; >=14 & <28=2; >=28=3 ???
  dplyr::mutate(EDAD_MUJER_REC= dplyr::case_when(EDAD_MUJER<18~"1.<18", EDAD_MUJER<17 & EDAD_MUJER<25~"2.18-24", EDAD_MUJER<24 & EDAD_MUJER<=35~"3.25-35",EDAD_MUJER>35~"4.>35"), HITO1_EDAD_GESTACIONAL_SEMANAS= dplyr::case_when(HITO1_EDAD_GESTACIONAL_SEMANAS>=4 & HITO1_EDAD_GESTACIONAL_SEMANAS<14~"1.4-13 semanas", HITO1_EDAD_GESTACIONAL_SEMANAS>=14 & HITO1_EDAD_GESTACIONAL_SEMANAS<28~"2.14-27 semanas", HITO1_EDAD_GESTACIONAL_SEMANAS>=28~ "3.>28 semanas"))%>% 
  dplyr::mutate(across(preds, ~ as.numeric(factor(.))+1))%>% dplyr::mutate(across(preds, ~ dplyr::case_when(is.na(.)~ 1, T~ .))) 

#comprobar 
#table(mydata_preds$sus_ini_mod_mvv, data2$sus_ini_mod_mvv, exclude=NULL)

f_preds <- item(AÑO, CAUSAL, EDAD_MUJER_REC, PAIS_ORIGEN_REC,HITO1_EDAD_GEST_SEM_REC, MACROZONA, PREV_TRAMO_REC) ~ 1

```

## Exploración

Modificamos la base de datos para incluir la variable resultado.

```{r, echo=FALSE}
#dropped region name because it was too different,many categories
preds2 <-c(setdiff(preds,""),"HITO2_DECISION_MUJER_IVE")  

mydata_preds2 <- data2%>% 
dplyr::mutate(across(preds, ~ as.numeric(factor(.))+1))%>% dplyr::mutate(across(preds, ~ dplyr::case_when(is.na(.)~ 1, T~ .)))  %>% dplyr::select(any_of(preds2)) %>% dplyr::mutate(outcome= factor(as.numeric(grepl('INTERRUMPIR', HITO2_DECISION_MUJER_IVE)))) 

#data.table::data.table(.)
#
#table(mydata_preds2$HITO2_DECISION_MUJER_IVE)
#lapply(preds2, function(p) {prop.table(table(mydata_preds2[p]))})
```

## Recodificación datos

Definimos los datos

```{r, echo=FALSE}
mydata_preds3 <-
  mydata_preds2 %>% 
#dplyr::mutate_if(is.numeric, ~as.character(.)) %>% #si convierto a caracter entrega errores
  data.table::data.table(.)
```

Así se ven los datos como los usa `glca`

Corremos `glca`.

## Análisis

```{r, glca-get, echo=FALSE}
#Biemer, P. P., & Wiesen, C. (2002). Measurement error evaluation of self-reported drug use: a latent class analysis of the US National Household Survey on Drug Abuse. Journal of the Royal Statistical Society: Series A (Statistics in Society), 165(1), 97–119. doi:10.1111/1467-985x.00612  
#lca_entropia(x="ppio", seed= 2125, k= 8, f= f_preds, dat= mydata_preds, nbr_repet= 30, na_rm= T)
#3
#<div style="border: 1px solid #ddd; padding: 5px; overflow-y: scroll; height:400px; overflow-x: scroll; width:100%">
# f is the selected variables
# dat is the data
# nb_var is the number of selected variables
# k is the number of latent class generated
# nbr_repet is the number of repetition to  
# reach the convergence of EM algorithm
# x es el código para las variables de los modelos
#seed es el numero random para las semillas. ej: 4345.
#Modo de calcular el mejor modelo.
#z_ # 
#2023-01-20
#https://github.com/QMUL/poLCAParallel/blob/master/exec/3_blrt.R
#0h s
seed<-2125
old <- Sys.time()

lca2 <- glca(f_preds, data = mydata_preds3, nclass = 2, seed = seed, verbose = FALSE, n.init = 5e2, decreasing=T, maxiter = 1e4,testiter = 500)
lca3 <- glca(f_preds, data = mydata_preds3, nclass = 3, seed = seed, verbose = FALSE, n.init = 5e2, decreasing=T, maxiter = 1e4,testiter = 500)
lca4 <- glca(f_preds, data = mydata_preds3, nclass = 4, seed = seed, verbose = FALSE, n.init = 5e2, decreasing=T, maxiter = 1e4,testiter = 500)
lca5 <- glca(f_preds, data = mydata_preds3, nclass = 5, seed = seed, verbose = FALSE, n.init = 5e2, decreasing=T, maxiter = 1e4,testiter = 500)
lca6 <- glca(f_preds, data = mydata_preds3, nclass = 6, seed = seed, verbose = FALSE, n.init = 5e2, decreasing=T, maxiter = 1e4,testiter = 500)
lca7 <- glca(f_preds, data = mydata_preds3, nclass = 7, seed = seed, verbose = FALSE, n.init = 5e2, decreasing=T, maxiter = 1e4,testiter = 500)
lca8 <- glca(f_preds, data = mydata_preds3, nclass = 8, seed = seed, verbose = FALSE, n.init = 5e2, decreasing=T, maxiter = 1e4,testiter = 500)
lca9 <- glca(f_preds, data = mydata_preds3, nclass = 9, seed = seed, verbose = FALSE, n.init = 5e2, decreasing=T, maxiter = 1e4,testiter = 500)
lca10 <- glca(f_preds, data = mydata_preds3, nclass = 10, seed = seed, verbose = FALSE, n.init = 5e2, decreasing=T, maxiter = 1e4,testiter = 500)

new_med<-(Sys.time())
paste0("The model took ",round(new_med-old,2)," until every LCA was computed")
```

Luego calculamos la razón de verosimilitud mediante remuestreo bootstrap (BLRT) entre los distintos modelos con el que asume una clase menos.

```{r,blrt, echo=FALSE}
gof<-
gofglca(lca2, lca3, lca4, lca5, lca6, lca7, lca8, lca9, lca10, test = "chisq")

bootlrt<-
gofglca(lca2, lca3, lca4, lca5, lca6, lca7, lca8, lca9, lca10, test = "boot", nboot=n_bootstrap, seed=2125)
                     
```

## Resultados

Hicimos un gráfico de los resultados

```{r, echo=FALSE}


manualcolors <- c('indianred1', 'cornflowerblue', 'gray50', 'darkolivegreen4', 'slateblue2', 
                  'firebrick4', 'goldenrod4')
levels3 <- c("logLik", "Gsq", "AIC", "CAIC", "BIC", "entropy", "Res.Df")
labels3 <- c('Log-Verosimilitud', 'Chi2', 'Criterio de Información\nde Akaike(AIC)', 'AIC Corregido', 'Criterio de Información\nBayesiano (BIC)', 'Entropía', 'Grados de libertad residuales')
fig_lca_fit3<- cbind.data.frame(rn=2:10,gof$gtable) %>%
  data.frame() %>% 
  dplyr::mutate_if(is.character, as.numeric) %>%  # convert character columns to numeric
  tidyr::pivot_longer(cols = -rn,names_to = "indices", values_to = "value", values_drop_na = F) %>%
  dplyr::mutate(indices = factor(indices, levels = levels3, labels = labels3)) %>%
  dplyr::filter(grepl("(AIC|BIC)",indices, ignore.case=T))%>%
	dplyr::mutate(ModelIndex= factor(rn, levels=2:10)) %>% 
  ggplot(aes(x = ModelIndex, y = value, group = indices, color = indices, linetype = indices)) +
  geom_line(size = 1.5) +
  scale_color_manual(values = manualcolors) +
  #scale_linetype_manual(values = c("solid", "dashed", "dotted")) +
  labs(x = "Número de clases", y = "Valor", color = "Medida", linetype = "Medida") +
  #facet_wrap(.~indices, scales = "free_y", nrow = 4, ncol = 1) +
  theme_bw()

fig_lca_fit3
ggsave("_fig1_comparison_glca_sin_po.png",fig_lca_fit3, dpi=600)
#International Journal of Workplace Health Management  (Zhang et al., 2018).
```

Luego en una tabla

::: controlly2
```{r, echo=FALSE}
cbind.data.frame(rn=2:10,gof$gtable) %>%#
  dplyr::select(rn, everything()) %>% 
    dplyr::mutate_if(is.character, as.numeric) %>%  # convert character columns to numeric
    knitr::kable(format="markdown", caption="Índices de ajuste modelos")
best_model<-
as.numeric(cbind.data.frame(rn=2:10,gof$gtable) %>% dplyr::summarise(which.min(BIC)+1))
```
:::

Presentamos el modelo con mejor ajuste

::: controlly
```{r, echo=FALSE}
summary(eval(parse(text = paste0("lca",best_model)))) #
```
:::

```{r, echo=FALSE}
save.image("data2_lca211_sin_po_alt.RData")
```

```{r, echo=FALSE}
require(tidyverse)
sesion_info <- devtools::session_info()
dplyr::select(
  tibble::as_tibble(sesion_info$packages),
  c(package, loadedversion, source)
) %>% 
  DT::datatable(filter = 'top', colnames = c('Row number' =1,'Variable' = 2, 'Percentage'= 3),
              caption = htmltools::tags$caption(
        style = 'caption-side: top; text-align: left;',
        '', htmltools::em('Packages')),
      options=list(
initComplete = htmlwidgets::JS(
        "function(settings, json) {",
        "$(this.api().tables().body()).css({
            'font-family': 'Helvetica Neue',
            'font-size': '50%', 
            'code-inline-font-size': '15%', 
            'white-space': 'nowrap',
            'line-height': '0.75em',
            'min-height': '0.5em'
            });",#;
        "}")))
```

