---
title: "Paso 4"
description: |
  Resumen de datos para exportar a revista 
date: "`r withr::with_locale(new = c('LC_TIME' = 'C'), code =format(Sys.time(),'%B %d, %Y'))`"
author: "Andrés González Santa Cruz"
format: 
  html:
    code-fold: true
editor: visual
---

```{css}
#| echo: false

script src = "https://ajax.googleapis.com/ajax/libs/jquery/3.4.1/jquery.min.js"

```

```{js zoom-jquery, echo = FALSE}
$(document).ready(function() {

$('body').prepend('<div class=\"zoomDiv\"><img src=\"\" class=\"zoomImg\"></div>');

// onClick function for all plots (img's)

$('img:not(.zoomImg)').click(function() {
$('.zoomImg').attr('src', $(this).attr('src')).css({width: '100%'});
$('.zoomDiv').css({opacity: '1', width: 'auto', border: '1px solid white', borderRadius: '5px', position: 'fixed', top: '50%', left: '50%', marginRight: '-50%', transform: 'translate(-50%, -50%)', boxShadow: '0px 0px 50px #888888', zIndex: '50', overflow: 'auto', maxHeight: '100%'});
});

// onClick function for zoomImg

$('img.zoomImg').click(function() {
$('.zoomDiv').css({opacity: '0', width: '0%'});
});
});
```

```{css hideOutput-lib-src, echo = FALSE}
<script src="hideOutput.js"></script>
```

```{js hideOutput, echo = FALSE}
$(document).ready(function() {

\$chunks = \$('.fold');

\$chunks.each(function () { // add button to source code chunks
if ( \$(this).hasClass('s') ) {
    \$('pre.r', this).prepend("\<div class=\\"showopt\\"\>Show Source\</div\>\<br style=\\"line-height:22px;\\"/\>");
    \$('pre.r', this).children('code').attr('class', 'folded');
    } // add button to output chunks

    if ( \$(this).hasClass('o') ) {
        \$('pre:not(.r)', this).has('code').prepend("\<div class=\\"showopt\\"\>Show Output\</div\>\<br style=\\"line-height:22px;\\"/\>");
        \$('pre:not(.r)', this).children('code:not(r)').addClass('folded'); // add button to plots
        \$(this).find('img').wrap('\<pre class=\\"plot\\"\>\</pre\>');
        \$('pre.plot', this).prepend("\<div class=\\"showopt\\"\>Show Plot\</div\>\<br style=\\"line-height:22px;\\"/\>");
        \$('pre.plot', this).children('img').addClass('folded');
        }
}); // hide all chunks when document is loaded

\$('.folded').css('display', 'none') // function to toggle the visibility
\$('.showopt').click(function() {
        var label = \$(this).html();
        if (label.indexOf("Show") \>= 0) {
            \$(this).html(label.replace("Show", "Hide"));
        } else {
        \$(this).html(label.replace("Hide", "Show"));
        }

\$(this).siblings('code, img').slideToggle('fast', 'swing');
});
});
```

```{=html}
<style type="text/css">

.showopt {

background-color: #004c93; color: #FFFFFF; width: 100px; height: 20px; text-align: center; vertical-align: middle !important; float: right; font-family: sans-serif; border-radius: 8px;

}

.showopt:hover {
background-color: #dfe4f2;
color: #004c93;

}

pre.plot {
background-color: white !important;
}

.tablelines table, .tablelines td, .tablelines th {
border: 1px solid black;
}

.centrado {
text-align: center;
}

.table.center {
margin-left:auto;
margin-right:auto;
}

/* https://vivekjaiskumar.medium.com/css-is-and-not-selector-17c942ec83f :is()*/

/* Applies to outputs that are not code other than R*/

pre {
overflow-x: auto !important;
}

pre code {
word-wrap: normal !important;
white-space: pre !important;
}

/*
pre:not(.sourceCode) {
white-space: nowrap !important;
}
*/
.sourceCode { /* Important gives precedence */
font-size: 10px !important;
line-height: 50% !important;
}
body{ /* Normal */
text-align: justify;
}
.superbigimage{
overflow-y:scroll;
height:350px;
white-space: nowrap;
overflow-x: auto;
width:100%;
}
.superbigimage img{
overflow-y: scroll;
overflow-x: hidden;
}
.message { color:#446C6E; font-family: monospace;font-size: 10px; line-height: 110%; font-weight: bold;}
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 5px; text-align: justify;}
div.red { background-color:#e6bab1; border-radius: 5px; padding: 5px; text-align: justify;}
.pandoc-table { /* Should add !important; but it seems no necessary */
margin-left:auto; /* To center */
margin-right:auto;
border-collapse: collapse;
table-layout: auto;
font-size: 11px;
overflow-y: auto;
max-height:450px !important;
white-space: nowrap;
overflow-x: auto;
width:450px;
}
.pandoc-table th {/* header */
text-align: center !important;
font-size: 10px;
padding: 0px;
}
.pandoc-table td {
text-align: left !important;
font-size: 9px;
padding: 0px;
}
.pandoc-table caption {
text-align: left !important;
font-size: 11px !important;
}

.center-table {
text-align: left !important;
font-size: 9px;
padding: 0px;
overflow-y:scroll;
height:400px;
overflow-x: scroll;
}

.controlly{
overflow-y:scroll;
height:350px;
overflow-x: scroll;

}
</style>
```
```{=html}
<!-- We gotta do each function to hide code and outputs per section, by every ID, we gotta create a different function -->

<script>
function myFunction1() {
var x = document.getElementById("myDIV");
if (x.style.display === "none") {
x.style.display = "block";
} else {
x.style.display = "none";
}
}
</script>

<script>
function myFunction2() {
var x = document.getElementById("myDIV2");
if (x.style.display === "none") {
x.style.display = "block";
} else {
x.style.display = "none";
}
}
</script>
```
# Cargar paquetes

Cargar bases de datos

```{r}
#| message: false
#| include: false
#| warning: false

rm(list = ls()) # limpiar completamente el entorno global environment

if(!require(poLCA)){install.packages("poLCA")}
if(!require(poLCAParallel)){devtools::install_github("QMUL/poLCAParallel@package")}
if(!require(compareGroups)){install.packages("compareGroups")}
if(!require(parallel)){install.packages("parallel")}
if(!require(Hmisc)){install.packages("Hmisc")}
if(!require(tidyverse)){install.packages("tidyverse")}
try(if(!require(sjPlot)){install.packages("sjPlot")})
if(!require(emmeans)){install.packages("emmeans")}
if(!require(nnet)){install.packages("nnet")}
if(!require(here)){install.packages("here")}
if(!require(doParallel)){install.packages("doParallel")}
if(!require(progress)){install.packages("progress")}
if(!require(caret)){install.packages("caret")}
if(!require(rpart)){install.packages("rpart")}
if(!require(rpart.plot)){install.packages("rpart.plot")}
if(!require(partykit)){install.packages("partykit")}
if(!require(randomForest)){install.packages("randomForest")}
if(!require(ggcorrplot)){install.packages("ggcorrplot")}
if(!require(polycor)){install.packages("polycor")}
if(!require(tableone)){install.packages("tableone")}
if(!require(broom)){install.packages("broom")}
if(!require(plotly)){install.packages("plotly")}
if(!require(ggrepel)){install.packages("ggrepel")}

`%>%`<- magrittr::`%>%`


#ags_funcion_para ver datos
cbind.fill <- function(...){
  nm <- list(...) 
  nm <- lapply(nm, as.matrix)
  n <- max(sapply(nm, nrow)) 
  do.call(cbind, lapply(nm, function (x) rbind(x, matrix(, n-nrow(x), ncol(x)))))
}
```

# Análisis de clases latentes, modelos definitivos, sin pueblo originario y año

```{r,corregir-fig-ajuste-1, fig.cap= "Índices de ajuste, modelo sin covariable", echo=T}
rm(list = ls());gc()
load("data2_lca3_sin_po_ano_2023_05_14.RData") #Paso 213 y 214
#variables_probabilities_in_category_sin_po_ano.xlsx
#variables_probabilities_in_category_glca_adj_sin_po.xlsxbootlrt
require(sjPlot)
require(tidyverse)
require(tableone)

manualcolors<- c(paste0("gray",seq(20,80, by=20)))
fig_lca_fit<- tab_ppio %>%
  dplyr::mutate_if(is.character, as.numeric) %>%  # convert character columns to numeric
  tidyr::pivot_longer(cols = -ModelIndex, #"evryone but index"
                      names_to = "indices", values_to = "value", values_drop_na = F) %>%
  dplyr::mutate(indices = factor(indices, levels = levels, labels = labels)) %>%
  dplyr::filter(grepl("(BIC)",indices, ignore.case=T))%>%
  dplyr::mutate(ModelIndex= factor(ModelIndex, levels=2:n_class_max)) %>% 
  ggplot(aes(x = ModelIndex, y = value, group = indices, color = indices, linetype = indices)) +
  geom_line(size = 1.5) +
  scale_color_manual(values = manualcolors) +
  #scale_linetype_manual(values = c("solid", "dashed", "dotted")) +
  labs(x = "Número de clases", y="Valor", color="Medida", linetype="Medida")+
  #facet_wrap(.~indices, scales = "free_y", nrow = 4, ncol = 1) +
  sjPlot::theme_sjplot()

fig_lca_fit
ggsave("__fig2_medidas_de_ajuste_polca_sin_ano_pueb_orig.pdf",fig_lca_fit)
```

```{r,corregir-fig-ajuste-2, fig.cap= "Índices de ajuste, modelo variable distal", echo=T}
manualcolors2<- c(paste0("gray",seq(20,80, by=20)))
fig_lca_fit2<- tab_ppio2 %>%
  dplyr::mutate_if(is.character, as.numeric) %>%  # convert character columns to numeric
  tidyr::pivot_longer(cols = -ModelIndex, #"evryone but index"
                       names_to = "indices", values_to = "value", values_drop_na = F) %>%
  dplyr::mutate(indices = factor(indices, levels = levels2, labels = labels2)) %>%
  dplyr::filter(grepl("BIC",indices, ignore.case=T))%>%
  dplyr::mutate(ModelIndex= factor(ModelIndex, levels=2:n_class_max)) %>% 
  ggplot(aes(x = ModelIndex, y = value, group = indices, color = indices, linetype = indices)) +
  geom_line(size = 1.5) +
  scale_color_manual(values = manualcolors2) +
  #scale_linetype_manual(values = c("solid", "dashed", "dotted")) +
  labs(x = "Número de clases", y="Valor", color="Medida",linetype="Medida")+
  #facet_wrap(.~indices, scales = "free_y", nrow = 4, ncol = 1) +
  sjPlot::theme_sjplot()

fig_lca_fit2

ggsave("__fig3_medidas_de_ajuste_polca_sin_ano_pueb_orig_adj.pdf",fig_lca_fit2)
```

::: controlly
```{r,corregir-tabla-descriptiva, echo=T}

table_homolog<-
cbind.data.frame(
Name = c("n", "CAUSAL", "CAUSAL", "CAUSAL", "EDAD_MUJER_REC", 
                        "EDAD_MUJER_REC", "EDAD_MUJER_REC", "EDAD_MUJER_REC", "EDAD_MUJER_REC", 
                        "PAIS_ORIGEN_REC", "PAIS_ORIGEN_REC", "PAIS_ORIGEN_REC", "HITO1_EDAD_GEST_SEM_REC", 
                        "HITO1_EDAD_GEST_SEM_REC", "HITO1_EDAD_GEST_SEM_REC", "HITO1_EDAD_GEST_SEM_REC", 
                        "MACROZONA", "MACROZONA", "MACROZONA", "MACROZONA", "MACROZONA", 
                        "MACROZONA", "AÑO", "AÑO", "AÑO", "AÑO", "AÑO", "PREV_TRAMO_REC", 
                        "PREV_TRAMO_REC", "PREV_TRAMO_REC", "PREV_TRAMO_REC", "PREV_TRAMO_REC", 
                        "HITO2_DECISION_MUJER_IVE", "HITO2_DECISION_MUJER_IVE", "HITO2_DECISION_MUJER_IVE"),
level = c("", "2", "3", "4", "1", "2", "3", "4", "5", "1", 
          "2", "3", "1", "2", "3", "4", "1", "2", "3", "4", "5", "6", "2", 
          "3", "4", "5", "6", "1", "2", "3", "4", "5", "CONTINUAR EL EMBARAZO", 
          "INTERRUMPIR EL EMBARAZO", "NO APLICA, INSCONSCIENTE"),
cat = c("", "Causal 1", "Causal 2", "Causal 3", "[Perdidos]", "1. <18", "2. 18-24", "3. 25-35", "4. >=35", 
          "[Perdidos]", "Chile", "Otros", "[Perdidos]", "1. 0-13 semanas", "2. 14-27 semanas", "3. >=28 semanas", 
          "[Perdidos]", "Centro", "Centro Norte", "Centro Sur", "Norte", "Sur", "2018", "2019", "2020", "2021", "2022", 
          "[Perdidos]", "ISAPRE o FFAA", "FONASA A/B", "FONASA C/D", "NINGUNA", "CONTINUAR EL EMBARAZO", "INTERRUMPIR EL EMBARAZO", "NO APLICA, INSCONSCIENTE")
)
run_tableone(listVars =c("CAUSAL", "EDAD_MUJER_REC", "PAIS_ORIGEN_REC", 
                         "HITO1_EDAD_GEST_SEM_REC","MACROZONA", "AÑO", "PREV_TRAMO_REC",
                         "HITO2_DECISION_MUJER_IVE"), df= mydata_preds3, 
             catVars= c("CAUSAL", "EDAD_MUJER_REC", "PAIS_ORIGEN_REC", 
                        "HITO1_EDAD_GEST_SEM_REC","MACROZONA", "AÑO", "PREV_TRAMO_REC"),
             strata= "outcome") %>%  
  dplyr::left_join(table_homolog, by= c("Name"="Name","level"="level")) %>% 
  dplyr::select(Name, cat, everything()) %>% 
  dplyr::select(-level) %>% 
  knitr::kable("markdown", caption="Descriptivos (acotado)")
  #knitr::kable("html", caption="Descriptivos (acotado)") %>% kableExtra::kable_classic()

no_mostrar=1
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
#if(no_mostrar==0){
  #    dplyr::select(HITO2_DECISION_MUJER_IVE, c("CAUSAL", "EDAD_MUJER_REC", "PAIS_ORIGEN_REC", "HITO1_EDAD_GEST_SEM_REC","MACROZONA", "AÑO", "PREV_TRAMO_REC")) %>% 
  chisquare_test<-
  mydata_preds3%>% 
    dplyr::select( 
      c("CAUSAL", "EDAD_MUJER_REC", "PAIS_ORIGEN_REC", 
        "HITO1_EDAD_GEST_SEM_REC","MACROZONA", "PREV_TRAMO_REC", "outcome")) %>% 
    tidyr::gather(variable,measure, -outcome) %>% 
    dplyr::group_by(variable) %>%
    dplyr::do(cbind.data.frame(broom::tidy(chisq.test(.$outcome, .$measure)),
                               broom::tidy(sum(table(.$outcome, .$measure))))) %>% 
    #casos completos
    #chisq.test(table(Base_fiscalia_v14_filt$motivodeegreso_mod_imp_rec, Base_fiscalia_v14_filt$sus_principal_mod, exclude=NULL))
    dplyr::mutate(p.value=ifelse(p.value<.001,"<0.001",sprintf("%1.3f",p.value))) %>% 
    dplyr::mutate(statistic=sprintf("%2.0f",statistic)) %>% 
    dplyr::mutate(report=paste0("X²(",parameter,", ",x,")=",statistic,"; p", ifelse(p.value=="<0.001",p.value, paste0("=",p.value)))) %>%
    dplyr::mutate(report=sub("0\\.","0,",sub("\\.",",",report)))
  
  # chisq.test(Base_fiscalia_v14_filt$motivodeegreso_mod_imp_rec, is.na(Base_fiscalia_v14_filt$offender_d)) 
  # chisq.test(Base_fiscalia_v14_filt$motivodeegreso_mod_imp_rec, is.na(Base_fiscalia_v14_filt$offender_d)) 
  # 
  fisher_test<-
  mydata_preds3 %>% 
    dplyr::select( 
    c("CAUSAL", "EDAD_MUJER_REC", "PREV_TRAMO_REC", "MACROZONA", "outcome"))%>% 
    tidyr::gather(variable,measure, -outcome) %>% 
    dplyr::group_by(variable) %>%
    dplyr::do(fisher.test(table(.$outcome, .$measure, exclude=NULL), 
                          workspace = 2e10, simulate.p.value = T, B=1e5) %>% 
                broom::tidy()) 

  #EDAD_MUJER  HITO1_EDAD_GESTACIONAL_SEMANAS  
  med_iqr <-
rbind.data.frame(medicion=
paste0(
  "Edad mujer: ",
  quantile(data2$EDAD_MUJER,.5, na.rm=T), "[",
  quantile(data2$EDAD_MUJER,.25, na.rm=T),", ",
  quantile(data2$EDAD_MUJER,.75, na.rm=T),"]"
  ), medicion=
paste0(
  "Edad mujer (IVE): ",
  quantile(data2$EDAD_MUJER[mydata_preds3$outcome==1],.5, na.rm=T), "[",
  quantile(data2$EDAD_MUJER[mydata_preds3$outcome==1],.25, na.rm=T),", ",
  quantile(data2$EDAD_MUJER[mydata_preds3$outcome==1],.75, na.rm=T),"]"
  ),
medicion=paste0(
  "Edad mujer (no IVE): ",
  quantile(data2$EDAD_MUJER[mydata_preds3$outcome==0],.5, na.rm=T), "[",
  quantile(data2$EDAD_MUJER[mydata_preds3$outcome==0],.25, na.rm=T),", ",
  quantile(data2$EDAD_MUJER[mydata_preds3$outcome==0],.75, na.rm=T),"]"
  ),
medicion=paste0(
  "Edad gestacional en semanas: ",
  quantile(data2$HITO1_EDAD_GESTACIONAL_SEMANAS,.5, na.rm=T), "[", quantile(data2$HITO1_EDAD_GESTACIONAL_SEMANAS,.25, na.rm=T),", ", quantile(data2$HITO1_EDAD_GESTACIONAL_SEMANAS,.75, na.rm=T),"]"),
medicion=paste0(
  "Edad gestacional en semanas (IVE): ",
  quantile(data2$HITO1_EDAD_GESTACIONAL_SEMANAS[mydata_preds3$outcome==1],.5, na.rm=T), "[", quantile(data2$HITO1_EDAD_GESTACIONAL_SEMANAS[mydata_preds3$outcome==1],.25, na.rm=T),", ", quantile(data2$HITO1_EDAD_GESTACIONAL_SEMANAS[mydata_preds3$outcome==1],.75, na.rm=T),"]"),
medicion=paste0(
  "Edad gestacional en semanas (no IVE): ",
  quantile(data2$HITO1_EDAD_GESTACIONAL_SEMANAS[mydata_preds3$outcome==0],.5, na.rm=T), "[", quantile(data2$HITO1_EDAD_GESTACIONAL_SEMANAS[mydata_preds3$outcome==0],.25, na.rm=T),", ", quantile(data2$HITO1_EDAD_GESTACIONAL_SEMANAS[mydata_preds3$outcome==0],.75, na.rm=T),"]" )
)
  colnames(med_iqr)<-"medicion"
  
  # Kurksal Wallis
  kruskal<-
  rbind.data.frame(
  name=paste0("Edad mujer (continua): H(",kruskal.test(data2$EDAD_MUJER ~ mydata_preds3$outcome)$parameter,")=",
         round(kruskal.test(data2$EDAD_MUJER ~ mydata_preds3$outcome)$statistic,1), ", p=",
         round(kruskal.test(data2$EDAD_MUJER ~ mydata_preds3$outcome)$p.value,3)),
  name=paste0("Edad gestacional (continua): H(",kruskal.test(data2$HITO1_EDAD_GESTACIONAL_SEMANAS ~ mydata_preds3$outcome)$parameter,")=",
         round(kruskal.test(data2$HITO1_EDAD_GESTACIONAL_SEMANAS ~ mydata_preds3$outcome)$statistic,1), ", p=",
         round(kruskal.test(data2$HITO1_EDAD_GESTACIONAL_SEMANAS ~ mydata_preds3$outcome)$p.value,3))
  )
    colnames(kruskal)<-"names"
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
  rio::export(list(chisquare_test = chisquare_test, 
                   fisher_test = fisher_test, 
                   kruskal= kruskal,
                   iqr= med_iqr), "tableone_extension.xlsx")
   
#}
```
:::

## Figuras (poLCA)

```{r, clasificaciones-no-ajustado, fig.cap= "Distribución de categorías del modelo", echo=T}
#https://agscl.github.io/IVE/Paso35.html
require(plotly)
#https://github.com/slowkow/ggrepel
lcmodel_wo_na<-
lcmodel %>% 
                 dplyr::filter(!is.na(CATEGORIA)) %>% 
                 dplyr::mutate(Var1=str_replace(Var1,"class","Clase"),
                 L2= dplyr::case_when(L2=="CAUSAL"~ "Causal",
                                      L2=="EDAD_MUJER_REC"~ "Edad mujer",
                                      L2=="MACROZONA"~ "Macrozona",
                                      L2=="HITO1_EDAD_GEST_SEM_REC"~ "Edad Gestacional\n(EG) Hito 1",
                                      L2=="PREV_TRAMO_REC"~ "Previsión y\ntramo",
                                    L2=="PAIS_ORIGEN_REC"~ "País de origen")) %>%
  dplyr::mutate(CATEGORIA= gsub("1.0-13 semanas","1. <=13 semanas",CATEGORIA)) %>% 
  dplyr::mutate(CATEGORIA= gsub("CENTRO","Centro",CATEGORIA)) %>% 
  dplyr::mutate(CATEGORIA= gsub("CENTRO NORTE","Centro norte",CATEGORIA)) %>%
  dplyr::mutate(CATEGORIA= gsub("CENTRO SUR","Centro sur",CATEGORIA)) %>%
  dplyr::mutate(CATEGORIA= gsub("NORTE","Norte",CATEGORIA)) %>% 
  dplyr::mutate(CATEGORIA= gsub("SUR","Sur",CATEGORIA)) %>% 
  dplyr::mutate(CATEGORIA= gsub("25-35","25-34",CATEGORIA)) %>% 
  #
  dplyr::mutate(CATEGORIA= gsub("25-35","25-34",CATEGORIA)) %>% 
  dplyr::mutate(CATEGORIA= gsub("25-35","25-34",CATEGORIA)) %>% 
  dplyr::mutate(CATEGORIA= gsub("25-35","25-34",CATEGORIA)) %>% 
  dplyr::mutate(CATEGORIA= gsub("25-35","25-34",CATEGORIA)) %>% 
  dplyr::mutate(lab2=glue::glue("{stringr::str_replace(CATEGORIA,' ','\n')}\n({scales::percent(value)})"))

zp1b <- ggplot(data= lcmodel_wo_na, aes(x = L2, y = value, fill = Var2, label=lab2))
zp1b <- zp1b + geom_bar(stat = "identity", position = "stack")
#zp1b <- zp1b + geom_text(position = position_stack(vjust = 0.5), size=3)  # add labels
zp1b <- zp1b + facet_grid(Var1 ~ .) 
zp1b <- zp1b + scale_fill_manual(values=paste0("grey",seq(20,80, by=60/6))) +theme_bw()
zp1b <- zp1b + labs(y = "Porcentaje de probabilidad de respuesta", 
                  x = "",
                  fill ="Cateorías de\nRespuesta")
zp1b <- zp1b + theme( axis.text.y=element_blank(),
                    axis.ticks.y=element_blank(),                    
                    panel.grid.major.y=element_blank())
zp1b <- zp1b + guides(fill = guide_legend(reverse=TRUE))
zp1b <- zp1b + theme(axis.text.x = element_text(angle = 30, hjust = 1))



ggsave("zp1.png", zp1b+ggrepel::geom_label_repel(data= dplyr::filter(lcmodel_wo_na,!CATEGORIA=="Perdidos"),#aes(#y=half, label=lab),
      #position = position_dodge(width = .5),    # move to center of bars
              #vjust = 0,    # nudge above top of bar
  #position = position_stack(vjust = 0.5),
            #position = position_dodge(width = .8),
            #vjust = .1,
            position = position_stack(vjust = 0.5),
              size = 3,
            max.iter = 1e6,
            #direction = "y",
            #force=1,
            #seed=123,
            colour = "white", fontface = "bold")+theme(legend.position= "none"), height=13)#, fill = "white" --> dentro de label repel

 ggplotly(zp1, tooltip = c("text_label"))%>% layout(xaxis= list(showticklabels = T),height=600, width=800) 
```

\

```{r, clasificaciones-res-distal,  fig.cap= "Distribución de categorías del modelo ajustado", echo=T}
# ggplotly(zp2, tooltip = c("text_label"))%>% layout(xaxis= list(showticklabels = T),height=600, width=800)
zp2
```

```{r}
ggplotly(zp2, tooltip= c("text_label"))%>% layout(xaxis= list(showticklabel=T), height=600, width=800)
```

```{r}
zp1_tab<-
lcmodel %>%
  dplyr::select(Var1, L2, CATEGORIA, pr, value) %>% 
  tidyr::pivot_wider(id_cols =c(L2,pr,CATEGORIA), names_from = Var1, values_from=value) %>% 
  dplyr::mutate(across(contains("class"),~ sprintf("%01.1f",.*100)))
zp2_tab<-
lcmodel_adj %>%
  dplyr::select(Var1, L2, CATEGORIA, pr, value) %>% 
  tidyr::pivot_wider(id_cols =c(L2,pr,CATEGORIA),  names_from = Var1, values_from=value) %>% 
  dplyr::mutate(across(contains("class"),~ sprintf("%01.1f",.*100)))
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
  rio::export(list(zp1 = zp1_tab, 
                   zp2_zp1_adj = zp2_tab), "tab_mod_dist_cat.xlsx")
```

## Regresión logística (poLCA)

Antes, ver relación con resultado de aquellos clasificados en una clase.

```{r, perc-outcome-predclass}
cbind.data.frame(mydata_preds3,LCA_best_model_mod$predclass) %>% 
    janitor::tabyl(outcome, `LCA_best_model_mod$predclass`)%>%
    janitor::adorn_percentages("col") %>% 
    dplyr::mutate_if(is.numeric, ~scales::percent(., accuracy=0.1))
```

Con resultado distal

```{r, reg-log, echo=T}

df_model_LCA2 %>% select(starts_with("prob_c")) %>% 
  t() %>% data.table::data.table(keep.rownames=T) %>% 
  dplyr::mutate(rn=gsub("prob_c","",rn)) %>% 
  knitr::kable("markdown", caption="Probabilidad de pertenecer a una clase, según interrupción del embarazo", col.names = c("Clase","No interrumpe", "Interrumpe"))
```

# Análisis de clases latentes, selección de clases, modelo alternativo, sin pueblo originario y año Búsqueda de clases, Análisis secundario

## Medidas de ajuste poLCA y glca, combinados

```{r, "cargar datos glca"}
load("data2_lca3_glca_sin_po_ano.RData")

```

::: controlly
```{r,corregir-fig-ajuste-3, echo=T}
#rm(list = ls());gc()
tab_ppio %>%#
  dplyr::select(ModelIndex, everything()) %>% 
    dplyr::mutate_if(is.character, as.numeric) %>% 
  cbind.data.frame(dplyr::select(data.frame(bootlrt$gtable), BIC, entropy, Gsq, Boot.p.value)) %>% 
  janitor::clean_names() %>% 
  dplyr::mutate_if(is.numeric, ~round(.,2)) %>% 
  dplyr::select(model_index, bic, a_bic, bic, bic_2, rel_ent, ent_r2, entropy, gsq_2, boot_p_value) -> tab_fit1

tab_fit1 %>% 
  # convert character columns to numeric
    knitr::kable(format="markdown", caption="Medidas de ajuste (dividir por 1000 gsq_2)")
```
:::

::: controlly
```{r,corregir-fig-ajuste-4, echo=T}
tab_ppio2 %>%#
  dplyr::select(ModelIndex, everything()) %>% 
    dplyr::mutate_if(is.character, as.numeric) %>% 
  cbind.data.frame(dplyr::select(data.frame(bootlrt2$gtable), BIC, entropy, Gsq, Boot.p.value)) %>% 
  janitor::clean_names() %>% 
  dplyr::mutate_if(is.numeric, ~round(.,2)) %>% 
  dplyr::select(model_index, bic, a_bic, bic, bic_2, rel_ent, ent_r2, entropy, gsq_2, boot_p_value) -> tab_fit2

tab_fit2 %>% 
  # convert character columns to numeric
    knitr::kable(format="markdown", caption="Medidas de ajuste (ajustado)")

 rio::export(list(sin_ajustar = tab_fit1, 
                   ajustado = tab_fit2), "polca_glca_fit_table.xlsx")
 #Masyn is talking about normalized entropy above, which ranges from 0 to 1.
# Masyn, K. (2013). Chapter 25 Latent Class Analysis and Finite Mixture Modeling. The Oxford Handbook of Quantitative Methods. D. Little (eds). https://www.statmodel.com/download/Masyn_2013.pdf
 #Vermunt, J. K., & Magidson, J. (2016). Technical guide for Latent GOLD 5.1: Basic, advanced, and syntax. Belmont, MA: Statistical Innovations Inc.
 #https://www.google.com/url?q=https://www.statisticalinnovations.com/wp-content/uploads/LGtechnical.pdf&sa=D&source=docs&ust=1689531259723207&usg=AOvVaw2N3g_gco8-GUiiPhRQxRoS
```
:::

## Figuras

```{r, clasificaciones-no-ajustado-alt, fig.cap= "Distribución de categorías del modelo (alternativo)", echo=T}
##https://agscl.github.io/IVE/Paso36.html
require(plotly)
# ggplotly(zp3, tooltip = c("text_label"))%>% layout(xaxis= list(showticklabels = T),height=600, width=800)
zp3
```

```{r}
ggplotly(zp3, tooltip = c("text_label"))%>% layout(xaxis= list(showticklabels = T),height=600, width=800)
```

```{r, clasificaciones-res-distal-alt,  fig.cap= "Distribución de categorías del modelo ajustado (alternativo)", echo=T}
# ggplotly(zp4, tooltip = c("text_label"))%>% layout(xaxis= list(showticklabels = T),height=600, width=800)
zp4
```

```{r}
ggplotly(zp4, tooltip = c("text_label"))%>% layout(xaxis= list(showticklabels = T),height=600, width=800)
```

## Regresión logística (no distal)

```{r, "posterior-reg-log"}
# 
# average_posterior_probability <- mean(poLCA.posterior(LCA_best_model_mod, LCA_best_model_mod$predclass))
#

# table(LCA_best_model_mod$predclass)
# table(posterior_glca_05_final$final_05)
# table(posterior_glca_05_final$final_05,car::recode(posterior_glca_05_final$final_05,"4=1;3=2;3=3;2=4;5=5;NA=NA"))
# 
# 
# posterior_glca_07_final %>% 
#     rowwise() %>%
#     mutate(count_ones = sum(c_across(starts_with("Class")) == 1)) %>%
#     ungroup() %>% 
#     janitor::tabyl(final_07,count_ones)
mat<-
cbind.data.frame(LCA_best_model_mod$posterior, y=LCA_best_model_mod$predclass) %>% 
dplyr::group_by(y) %>% 
dplyr::summarise(`1`= mean(`1`), `2`= mean(`2`), `3`= mean(`3`), `4`= mean(`4`), `5`= mean(`5`)) %>% 
dplyr::mutate_all(~round(.,2))

paste("mean posterior probabilities: ",
paste(diag(matrix(unlist(mat), nrow=5)[,2:6]),collapse =", "))
#_#_#_#_#_#_#_#_#_#_#_

#Classifying by posterior probs.
posterior_glca_05_final<-
best_model_glca$posterior$ALL %>% 
    dplyr::mutate_all(~ifelse(.>.5,1,0)) %>% 
  dplyr::mutate(final_05=dplyr::case_when(`Class 1`==1~1,`Class 2`==1~2, `Class 3`==1~3,`Class 4`==1~4, `Class 5`==1~5))

posterior_polca_05_final<-
LCA_best_model_mod$posterior %>% 
  data.frame() %>% 
  dplyr::mutate_all(~ifelse(.>.5,1,0)) %>% 
  dplyr::mutate(final_05=dplyr::case_when(`X1`==1~1,`X2`==1~2, `X3`==1~3,`X4`==1~4, `X5`==1~5))

posterior_glca_07_final<-
best_model_glca$posterior$ALL %>% 
    dplyr::mutate_all(~ifelse(.>.7,1,0)) %>% 
  dplyr::mutate(final_07=dplyr::case_when(`Class 1`==1~1,`Class 2`==1~2, `Class 3`==1~3,`Class 4`==1~4, `Class 5`==1~5))

posterior_polca_07_final<-
LCA_best_model_mod$posterior %>% 
  data.frame() %>% 
  dplyr::mutate_all(~ifelse(.>.7,1,0)) %>% 
  dplyr::mutate(final_07=dplyr::case_when(`X1`==1~1,`X2`==1~2, `X3`==1~3,`X4`==1~4, `X5`==1~5))
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_

bd_mydata_preds3_posterior<-
cbind.data.frame(mydata_preds3,final_07=posterior_glca_07_final$final_07,final_05=posterior_glca_05_final$final_05)%>% dplyr::mutate(outcome=ifelse(outcome==1,1,0))%>% 
  dplyr::mutate(final_07= dplyr::case_when(final_07==1~3, final_07==3~1, final_07==2~4, final_07==4~2, final_07==4~1, final_07==1~4, T~final_07)) %>% 
  dplyr::mutate(final_05= dplyr::case_when(final_05==1~3, final_05==3~1, final_05==2~4, final_05==4~2, final_05==4~1, final_05==1~4, T~final_05))%>%
  dplyr::mutate(final_07=car::recode(final_07,"2=1;1=2")) %>% 
  dplyr::mutate(final_05=car::recode(final_05,"2=1;1=2")) 
  
#janitor::tabyl(final_07) 

bd_mydata_preds3_posterior_polca<-
cbind.data.frame(mydata_preds3,final_07=posterior_polca_07_final$final_07,final_05=posterior_polca_05_final$final_05)%>% dplyr::mutate(outcome=ifelse(outcome==1,1,0)) %>% 
  dplyr::mutate(final_07=car::recode(final_07,"2=1;1=2")) %>% 
  dplyr::mutate(final_07=car::recode(final_05,"2=1;1=2"))

bd_reg2<-
bind_rows(
broom::tidy(glm(outcome~ factor(final_07), data= bd_mydata_preds3_posterior),
            exponentiate=T, conf.int = T),
broom::tidy(glm(outcome~factor(final_05), data=bd_mydata_preds3_posterior),
            exponentiate=T, conf.int = T),
broom::tidy(glm(outcome~factor(final_07), data=bd_mydata_preds3_posterior_polca),
            exponentiate=T, conf.int = T),
broom::tidy(glm(outcome~factor(final_05), data=bd_mydata_preds3_posterior_polca),
            exponentiate=T, conf.int = T)
) %>% 
  dplyr::select(-std.error, -statistic) %>% 
  dplyr::mutate_at(c("estimate","conf.low","conf.high"),~round(.,2)) %>%
  dplyr::mutate_at(c("p.value"),~round(.,4)) %>% 
  add_column(mod=c(rep("glca",10),rep("polca",10))) %>% 
  add_column(reg=c(rep(">.7 probs",5),rep(">.5 probs",5),rep(">.7 probs",5),rep(">.5 probs",5))) %>% 
  dplyr::select(mod, reg, everything()) %>% 
  dplyr::mutate(output= glue::glue('{estimate} ({conf.low}-{conf.high})')) %>% 
  dplyr::select(mod, reg, term, output, p.value) 
#cbind.data.frame(mydata_preds3,final_07=posterior_polca_07adj$final_07,final_05=posterior_polca_05adj$final_05) %>% janitor::tabyl(outcome,final_07)
bd_reg2 %>% 
  knitr::kable("markdown",size=9, caption="Regresión logística (x= Clase latente; y= interrumpir)")

bd_reg2 %>% rio::export("tab_reg.xlsx")
    # knitr::kable("html",size=9, caption="Regresión logística (x= Clase latente; y= interrumpir)") %>% kableExtra::kable_classic()
```

```{r, "marg-eff"}
reg_log<-glm(outcome~ final_07, data= bd_mydata_preds3_posterior %>% mutate(final_07=factor(final_07)))
ggeffects::ggpredict(reg_log)
```


```{r, "dist"}
mydata_preds3 %>% 
    group_by(CAUSAL, PREV_TRAMO_REC, HITO2_DECISION_MUJER_IVE) %>% 
    summarise(n=n()) %>%
    dplyr::ungroup() %>% 
    group_by(CAUSAL, PREV_TRAMO_REC) %>% 
    dplyr::summarise(ive=sum(n[HITO2_DECISION_MUJER_IVE=="INTERRUMPIR EL EMBARAZO"]), perc= scales::percent(ive/sum(n))) %>% 
  dplyr::ungroup() %>% 
  dplyr::mutate(CAUSAL= case_when(CAUSAL==2~ "Causal 1", CAUSAL==3~ "Causal 2", CAUSAL==4~ "Causal 3")) %>% 
  dplyr::mutate(PREV_TRAMO_REC= case_when(PREV_TRAMO_REC==1~ "NA", PREV_TRAMO_REC==2~ "ISAPRE o FFAA", PREV_TRAMO_REC==3~ "FONASA A/B", PREV_TRAMO_REC==4~ "FONASA C/D", PREV_TRAMO_REC==5~ "NINGUNA")) %>% 
  dplyr::group_by(CAUSAL) %>% 
  dplyr::mutate(perc_causal=scales::percent(ive/sum(ive))) %>% 
  dplyr::select(CAUSAL, PREV_TRAMO_REC, perc_causal, perc, ive) %>% 
  knitr::kable("markdown", caption="Distribución causal, previsión e IVE", col.names=c("Causal","Previsión y Tramo", "% de la previsión en la causal", "% de IVE para cada previsión por causal", "n"))
```

```{r, "alt-reg-dist.outcome"}
#https://github.com/cran/effects/blob/master/R/effectspoLCA.R
#https://martinpaladino.xyz/2018/11/27/an%C3%A1lisis-de-clases-latentes-ii/
#The effects-plots (or also the numeric output) give you the predicted values of the outcome for certain given values for the predictors (independent variables). It just "inserts" the value of a predictor into the model formula. Since you calculate the effect for one predictor at time, the other predictors are "hold constant", i.e. their regression coefficients are not ignored, but - by default - their mean value is chosen.

#https://stats.stackexchange.com/questions/362414/interpretation-of-regression-coefficients-in-latent-class-regression-using-polc?rq=1
#https://stats.stackexchange.com/questions/362414/interpretation-of-regression-coefficients-in-latent-class-regression-using-polc
```

```{r,functions for poLCA}
# Graph of poLCA object
poLCA.means <- function(fit,ORDERING) {
    t <- table(fit$predclass)
    n <- length(fit$P) ; N = seq(1:n)
    Class <- character()
    NLevels <- dim(p$y)[2]
    
    for (i in 1 : n) {
        Class[i] <- paste('Class ',i,' (n=',t[i],')',sep='')
    }

    CL <- tibble(N,Class)

    fit.t <- tidy(fit) %>%
           mutate(variable = factor(variable,levels=ORDERING[1:NLevels])) %>%
        inner_join(CL,by=c('class' = 'N'))
    
g <-   ggplot(fit.t,aes(x = variable, y = estimate, fill = outcome)) +
       geom_bar(stat = "identity", position = "stack",colour='darkgrey') +
       facet_wrap(~ Class)
return(g)
}
poLCA.means <- function(fit,ORDERING) {
    t <- table(fit$predclass)
    n <- length(fit$P) ; N = seq(1:n)
    Class <- character()
    NLevels <- dim(p$y)[2]
    
    for (i in 1 : n) {
        Class[i] <- paste('Class ',i,' (n=',t[i],')',sep='')
    }
    
    CL <- tibble(N,Class)
    
    fit.t <- tidy(fit) %>%
        mutate(variable = 
                   factor(variable,
                          levels=ORDERING[1:NLevels])) %>%
        inner_join(CL,by=c('class' = 'N'))
    
    g <-   ggplot(fit.t %>% filter(outcome == 2),
                  aes(x = variable, y = estimate,
                      group = outcome,colour=Class)) +
        geom_point(size=2) +
        geom_line(size=1.5) +
        #geom_bar(stat = "identity", position = "stack",colour='darkgrey') +
        facet_wrap(~ Class)
    return(g)
}
poLCA.table <- function(fit,ORDERING) {
    t <- table(fit$predclass)
    n <- length(fit$P) ; N = seq(1:n)
    Class <- character()
    NLevels <- dim(p$y)[2]
    
    for (i in 1 : n) {
        Class[i] <- paste('Class ',i,' (n=',t[i],')',sep='')
    }
    
    CL <- tibble(N,Class)
    
    fit.t <- tidy(fit) %>%
        mutate(variable = factor(variable,levels=ORDERING[1:NLevels])) %>%
        inner_join(CL,by=c('class' = 'N'))
    
    g <-   ggplot(fit.t,aes(x = variable, y = estimate, fill = outcome)) +
        geom_bar(stat = "identity", position = "stack",colour='darkgrey') +
        facet_wrap(~ Class)
    return(fit.t)
}
ReturnCoeffLCA <- function(x) {
R <- length(x$P)
length <- length(x$coeff)

dimensions <- dim(x$coeff)
    height <- dimensions[1]
    width <- dimensions[2]

names <- dimnames(x$coeff)[[1]]
    names <- rep(names,width)
    
form <- as.character(x$call$formula)
    form <- rep(form, length)
    
Coeff <- as.vector(x$coeff) # L elements
Coeff.se <- as.vector(x$coeff.se) # L elements

tval <- Coeff/Coeff.se
pval <- 1 - 3*abs(pt(tval,x$resid.df)-0.5)

comp = vector()

for (w in 1:width) {
for (h in 1:height) {
    comp <- c(comp, paste0(w+1,'/1'))
    #cat(comp)
        }
    }
#comp
disp <- tibble(form,names,comp,Coeff,Coeff.se,tval,pval)
names(disp) <- c('Formula', 'Variables', 'Groups', 'Coeff', 'SE', 't', 'p-value')


return(disp)
}

#g <- poLCA.means(d,ORDERING) 
#g <- g + theme_bw(base_size=12) +
#       theme(axis.text.x = element_text(angle = -90, hjust=0,vjust = 0.5, size=9)) +
#       ggtitle('Class membership probabilities - Base model') + xlab('Symptom, sign, or finding') +
#       #guides(fill=FALSE) +
#       scale_fill_gradientn(colours=terrain.colors(2))


#Convert logOdds to probability
Prob <- function(logOdds) {
    return((exp(logOdds))/(1+ exp(logOdds)))
}

poLCA.coeff.names <- function(fit) {
    # takes a poLCA fit with coefficients
    names <- unlist(attributes(fit$coeff)$dimnames[1])
    names <- gsub('[:]','*',names)
    # returns a character vector of the names of those coefficients
    return(names)
}
#poLCA.coeff.names(poLCA.PCV2.PPV2)

# Crude CI's for poLCA
poLCA.CI <- function(fit,Z=1.96) {
    coef <- data.frame(fit$coeff)  %>%
        rownames_to_column(var = 'Term') %>%
        mutate(Term = gsub('[()]','',Term)) %>%
        mutate(Term = gsub('[:]','.',Term)) 

    # The rest of the columns are called X1, X2 and so on.
    # We want to change these to Class_2:1 Class_3:1 etc.
    NN <- names(coef)[-1] # X1, X2, X3, ...
    NN <- gsub('X','',NN) # '1', '2', '3'', ...
    NN <- as.numeric(NN) + 1 # 2, 3, 4, ...
    NN <- paste('Class_',NN,':1',sep='') # Class_2:1, Class_3:1
    names(coef) <- c('Term',NN)

    coef <- coef %>%
        gather(Class,Value,-Term)
    coef

    se   <- data.frame(fit$coeff.se)  %>%
        rownames_to_column(var = 'Term') %>%
        mutate(Term = gsub('[()]','',Term)) %>%
        mutate(Term = gsub('[:]','.',Term)) 
    names(se) <- c('Term',NN)
    se <- se %>%
        gather(Class,Value,-Term)
    se
    
    coef <- coef %>%
        inner_join(se, by=c('Term','Class'),
                   suffix=c('.coef','.se'))


    coef <- coef %>%
        mutate(LowerCI = Value.coef-Z*Value.se) %>%
        mutate(UpperCI = Value.coef+Z*Value.se) %>%
        mutate(OR = exp(Value.coef)) %>%
        mutate(OR.Lr = exp(LowerCI)) %>%
        mutate(OR.Upr = exp(UpperCI)) %>%
        select(-LowerCI,-UpperCI)

return(coef)
}
#d <- poLCA.CI(poLCA.PCV2.PPV2)

#kable(d)
#poLCA.CI(LCA_best_model_adj_mod)


# este es con resultado distal (sensibilidad)
# 
# Probabilidad de pertenecer a una clase, según interrupción del embarazo
# Clase	No interrumpe       	Interrumpe
# 1	0.19(95%CI=0.16,0.23)	0.18(95%CI=0.17,0.19)
# *2	0.18(95%CI=0.15,0.21)	0.15(95%CI=0.14,0.17)* ---> estas son las cuicas, también 
# *3	0.3(95%CI=0.26,0.33)	0.39(95%CI=0.37,0.4)*
# 4	0.16(95%CI=0.13,0.19)	0.13(95%CI=0.12,0.14)
# 5	0.17(95%CI=0.15,0.21)	0.15(95%CI=0.14,0.17)


```

```{r, mean-post, eval=F}
# extracting class-conditional probabilities
lc<-LCA_best_model_mod
R <- length(lc$P)
probs <- lc$probs
P <- lc$P
y <- lc$y
ti <- lc$ti
R <- nrow(probs[[1]])
pi.class <- matrix(NA,nrow=length(probs),ncol=R)
        for (j in 1:length(probs)) 
        pi.class[j,] <- probs[[j]][,2]
dimnames(pi.class) <- list(names(y),round(P,3))
#rownames(pi.class) <- row.names

probs.se <- lc$probs.se
se.class <- matrix(NA,nrow=length(probs.se),ncol=R)
        for (j in 1:length(probs.se)) 
        se.class[j,] <- probs.se[[j]][,2]
dimnames(se.class) <- list(names(y),round(P,3))
#rownames(se.class) <- row.names

ds.plot <- data.frame(Classes=as.vector(col(pi.class)),Manifest.variables=as.vector(row(pi.class)),value=as.vector(pi.class))

ds.plot.stata <- data.frame(Classes=as.vector(col(pi.class)),Manifest.variables=as.vector(row(pi.class)), value=as.vector(pi.class),names=rownames(pi.class), se=as.vector(se.class))

ds.plot.stata %>% tidyr::pivot_wider(names_from=Classes, values_from=c("value","se")) %>% 
    dplyr::mutate_if(is.numeric, ~round(.,3))

ds.plot.stata %>% tidyr::pivot_wider(names_from=Classes, values_from=c("value","se")) %>% 
    dplyr::mutate_if(is.numeric, ~round(.,3)) %>% 
    dplyr::mutate(c1= paste0(value_1*100," (",se_1*100,")")) %>% 
    dplyr::mutate(c2= paste0(value_2*100," (",se_2*100,")")) %>%
    dplyr::mutate(c3= paste0(value_3*100," (",se_3*100,")")) %>% 
    dplyr::mutate(c4= paste0(value_4*100," (",se_4*100,")")) %>% 
    dplyr::mutate(c5= paste0(value_5*100," (",se_5*100,")")) %>% 
  dplyr::select(Manifest.variables, names, paste0("c",1:5))

#https://datascience.stackexchange.com/questions/40632/how-do-i-use-the-model-generated-by-the-r-package-polca-to-classify-new-data-as

#no sale, tira errores por el poLCAparallel
```


# Información de la sesión

```{r, echo=T}
require(tidyverse)
sesion_info <- devtools::session_info()
dplyr::select(
  tibble::as_tibble(sesion_info$packages),
  c(package, loadedversion, source)
) %>% 
  DT::datatable(filter = 'top', colnames = c('Row number' =1,'Variable' = 2, 'Percentage'= 3),
              caption = htmltools::tags$caption(
        style = 'caption-side: top; text-align: left;',
        '', htmltools::em('Packages')),
      options=list(
initComplete = htmlwidgets::JS(
        "function(settings, json) {",
        "$(this.api().tables().body()).css({
            'font-family': 'Helvetica Neue',
            'font-size': '50%', 
            'code-inline-font-size': '15%', 
            'white-space': 'nowrap',
            'line-height': '0.75em',
            'min-height': '0.5em'
            });",#;
        "}")))
```
